---
title: 'Exploratory Data Ananlysis: Low-Income Energy Affordability Data'
author: "Eric Scheier"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  md_document:
    variant: markdown_github
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(MASS)
library(scales)
library(caret)
library(tidyverse)
library(spatstat)
library(GGally)
library(ggpubr)
library(patchwork)
library(mgcv)
```

# Questions:

+ How many ACS survey responses inform each segment?
+ Are there any trends in the energy burden based on age, units/building, home ownership, home attachment, ?
+ Perform a principal component analysis
+ Does `year_constructed` or `min_age` better explain the energy burden variance?
+ Does `number_of_units` or `min_units` better explain the energy burden variance?

```{r}
data <- read_csv(file.path(getwd(),"clean_lead.csv"),
                          col_types=cols(
                            geo_id = readr::col_character(),
                            state_id = readr::col_character(),
                            county_id = readr::col_character(),
                            tract_id = readr::col_character(),
                            puma10_code = readr::col_character(),
                            fmr_code = readr::col_character(),
                            occupancy_type = readr::col_factor(),
                            income_bracket = readr::col_factor(),
                            primary_heating_fuel = readr::col_factor(),
                            number_of_units = readr::col_factor(),
                            year_constructed = readr::col_factor(),
                            households = col_double(),
                            annual_income = col_double(),
                            electricity_spend = col_double(),
                            gas_spend = col_double(),
                            other_spend = col_double(),
                            acs_responses = col_double(),
                            min_age = col_double(),
                            min_units = col_double(),
                            detached = readr::col_factor(NULL),
                            mean_energy_cost = col_double(),
                            mean_energy_burden = col_double(),
                            mean_eroi = col_double(),
                            energy_poverty = col_logical(),
                            pct_non_male = col_double(), 
                            median_age = col_double(),
                            pct_no_high_school = col_double(),
                            pct_non_white = col_double(),
                            median_income = col_double(),
                            acs_total_households = col_double(),
                            utility_id = readr::col_factor(levels = NULL, ordered = FALSE, include_na = TRUE),
                            percent_tract_served = readr::col_double(),
                            tract_households_served = readr::col_double(),
                            utility_name = readr::col_character(),
                            ownership_type = readr::col_factor(NULL),
                            balancing_authority = readr::col_factor(NULL),
                            retail_revenue_thousands = readr::col_double(),
                            retail_sales_MWh = readr::col_double(),
                            retail_customers = readr::col_double(),
                            implied_price_kWh = readr::col_double(),
                            revenue_per_customer = readr::col_double(),
                            implied_electricity_use = readr::col_double()
                            ))

summary(data)
```


```{r}
str(data)
```



```{r}
clean_data <- drop_na(data)
summary(clean_data)
```


## How many ACS survey responses inform each segment?

```{r}
rph <- data %>% 
  mutate(households_per_response=households/acs_responses) %>% 
  dplyr::select(households_per_response, 
         households,
         acs_responses)
  
summary(rph)
```

```{r}
ggplot(data=rph, aes(x=households,y=acs_responses)) + 
  geom_point() + 
  stat_smooth(method=lm)
```

```{r}
rph %>% 
  arrange(desc(households_per_response)) %>% 
  slice(1:10)
```

```{r}
rph %>% 
  arrange((households_per_response)) %>% 
  slice(1:10)
```

```{r eval=FALSE}
rph %>% 
  ggplot(aes(x=households_per_response, y=households)) + 
  geom_density(alpha=.2, fill="#FF6666", na.rm=TRUE) + 
  geom_histogram(binwidth=1, stat="identity") + 
  scale_x_continuous(limits = c(NA, 100))
```


```{r}
data %>% group_by(tract_id) %>% summarise(lead_households=sum(households), acs_households=mean(acs_total_households)) %>% 
  ggplot(aes(x=lead_households, y=acs_households)) + 
  geom_point() + stat_smooth(method=lm)
```

```{r}
weighted_means <- clean_data %>%# drop_na() %>% 
  group_by(ownership_type) %>% 
  summarise(households_sum = sum(households),
            cost_mean = weighted.mean(mean_energy_cost, households),
            electric_spend_mean = weighted.mean(electricity_spend, households),
            electric_use_mean = weighted.mean(implied_electricity_use, households),
            electric_cost_mean = weighted.mean(implied_price_kWh, households),
            energy_burden_mean = weighted.mean(mean_energy_burden, households),
            mean_income = weighted.mean(annual_income, households),
            mean_eroi = weighted.mean(mean_eroi, households))
weighted_means
```


```{r}
weighted_medians <- clean_data %>%# drop_na() %>% 
   group_by(ownership_type) %>% 
   summarise(median_electricity_spend = if( sum(!is.na(households))<3 ){NA} else { weighted.median(electricity_spend, households, na.rm=TRUE)},
             median_eroi = if( sum(!is.na(households))<3 ){NA} else { weighted.median(mean_eroi, households, na.rm=TRUE)},
            median_income = if( sum(!is.na(households))<3 ){NA} else { weighted.median(annual_income, households, na.rm=TRUE)},
            median_electricity_price = if( sum(!is.na(households))<3 ){NA} else { weighted.median(implied_price_kWh, households, na.rm=TRUE)},
            median_electricity_use = if( sum(!is.na(households))<3 ){NA} else { weighted.median(implied_electricity_use, households, na.rm=TRUE)},
            median_energy_burden = if( sum(!is.na(households))<3 ){NA} else { weighted.median(mean_energy_burden, households, na.rm=TRUE)},
            median_energy_cost = if( sum(!is.na(households))<3 ){NA} else { weighted.median(mean_energy_cost, households, na.rm=TRUE)}
            )

weighted_medians
```

```{r}
#clean_data$ownership_type_household_weights

clean_data <- clean_data %>% 
  group_by(ownership_type) %>% 
  mutate(ownership_households = sum(households)) %>% 
  mutate(ownership_type_household_weights = households/ownership_households)
```


```{r}
mean_eroi <- clean_data %>% #drop_na() %>%
  ggplot(aes(x=mean_eroi, weight=ownership_type_household_weights/3, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Household Energy Return on Investment By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::unit_format(unit = "x", scale = 1, accuracy = 1),
                     breaks=seq(from=0,to=100,by=10), 
                     minor_breaks=seq(from=0,to=100,by=1),
                     limits=c(0,100), name="Household Energy Return on Investment\n(Income Earned for each Dollar Spent on Energy)") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.05), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_eroi,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75) + 
  geom_vline(xintercept = 10, linetype="dotted", 
                color = "red", size=1.0, alpha=0.75) + 
  annotate("text", x = 10, y = 0.0075/3, angle = 90, color="red", label = "Energy Poverty Line", 
    vjust = -0.5, parse = FALSE, alpha=0.75) + 
  annotate("text", x = min(weighted_medians$median_eroi), y = 0.005/3, angle = 90, color="gray25", label = "Median", 
    vjust = -0.5, parse = FALSE, alpha=0.75)

mean_eroi
```

```{r}
federal_poverty_line <- 16240
annual_income <- clean_data %>% #drop_na() %>%
  ggplot(aes(x=annual_income, weight=ownership_type_household_weights/3, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Household Annual Income By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::dollar_format(), limits=c(0,150000), name="Annual Income") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.0005), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_income,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75) +
  geom_vline(xintercept = federal_poverty_line, linetype="dotted",
                color = "red", size=1.0, alpha=0.75)

annual_income +
  annotate("text", x = federal_poverty_line, y = 0.0000035/3, angle = 90, color="red", label = "Federal Poverty Line",
    vjust = -0.5, parse = FALSE, alpha=0.75) +
  annotate("text", x = min(weighted_medians$median_income), y = 0.0000025/3, angle = 90, color="gray25", label = "Median",
    vjust = -0.5, parse = FALSE, alpha=0.75)
```


```{r}
mean_energy_cost <- clean_data %>%
  ggplot(aes(x=mean_energy_cost, weight=ownership_type_household_weights/3, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Household Monthly Energy Spending By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::dollar_format(), limits=c(0,300), name="Energy Spend",
                     breaks=seq(from=0,to=300,by=50), 
                     minor_breaks=seq(from=0,to=300,by=10)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.0005), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_energy_cost,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75)

mean_energy_cost + annotate("text", x = min(weighted_medians$median_energy_cost), y = 0.0025/3, angle = 90, color="gray25", label = "Median",vjust = -0.5, parse = FALSE, alpha=0.75)
```

```{r}
electricity_spend <- clean_data %>%
  ggplot(aes(x=electricity_spend, weight=ownership_type_household_weights/3, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Household Monthly Electricity Spending By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::dollar_format(), limits=c(0,300), name="Energy Spend",
                     breaks=seq(from=0,to=300,by=50), 
                     minor_breaks=seq(from=0,to=300,by=10)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.0005), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_electricity_spend,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75)

electricity_spend + annotate("text", x = min(weighted_medians$median_electricity_spend), y = 0.0025/3, angle = 90, color="gray25", label = "Median",vjust = -0.5, parse = FALSE, alpha=0.75)
```

```{r}
implied_price_kWh <- clean_data %>%
  ggplot(aes(x=implied_price_kWh, weight=ownership_type_household_weights/3000, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Electricity Price By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::dollar_format(), limits=c(0.1,0.15), name="Electricity Unit Cost ($/kWh)",
                     breaks=seq(from=0.1,to=0.15,by=0.01), 
                     minor_breaks=seq(from=0.1,to=0.15,by=0.0025)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_electricity_price,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75) +
  annotate("text", x = min(weighted_medians$median_electricity_price), y = 0.1, angle = 90, color="gray25", label = "Median",vjust = -0.5, parse = FALSE, alpha=0.75)

implied_price_kWh
```

```{r}
implied_electricity_use <- clean_data %>%
  ggplot(aes(x=implied_electricity_use, weight=ownership_type_household_weights/3, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Implied Monthly Household Electricity Use By Utility Ownership Type") + 
  scale_x_continuous(labels = comma, limits=c(0,2200), 
                     name="Electricity Consumption (kWh)",
                     breaks=seq(from=0,to=2200,by=500), 
                     minor_breaks=seq(from=0,to=2200,by=100)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.0005), 
                     name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_electricity_use,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75) +
  annotate("text", x = min(weighted_medians$median_electricity_use), y = 0.00015, angle = 90, color="gray25", label = "Median",vjust = -0.5, parse = FALSE, alpha=0.75)

implied_electricity_use
```


```{r}
contour_plot <- ggplot(clean_data, aes(x=annual_income,
                                       y=mean_energy_cost,
                                       weight=ownership_type_household_weights/3,
                                       color=ownership_type,
                                       fill=ownership_type)) +
  stat_density_2d(aes(alpha = ..piece..), geom="polygon") +
  guides(alpha = FALSE) +
  stat_smooth(method = "lm", fullrange = TRUE) +
  geom_rug() + 
  scale_y_continuous(name = "Monthly Energy Cost", 
                     labels = scales::dollar_format(),
                     limits = c(0, 300), expand = c(0, 0)) + 
  scale_x_continuous(name = "Annual Income", 
                     labels = scales::dollar_format(),
                     limits = c(0, 150000), expand = c(0, 0)) + 
  theme_pubr() + #   theme(plot.margin = margin()) + 
  theme(legend.position = "bottom")

contour_plot# + facet_grid(. ~ ownership_type)
```


```{r}
dens1 <- ggplot(clean_data, aes(x = annual_income, weight=ownership_type_household_weights/3, 
                                fill = ownership_type)) + 
  geom_density(alpha = 0.4) + 
  theme_void() + xlim(0,250000) + 
  theme(legend.position = "none")

dens2 <- ggplot(clean_data, aes(x = mean_energy_cost, weight=ownership_type_household_weights/3, 
                                fill = ownership_type)) + 
  geom_density(alpha = 0.4) + 
  theme_void() + xlim(1,300) + 
  theme(legend.position = "none") + 
  coord_flip()
```


```{r}
# https://stackoverflow.com/a/59581217/2668097
annual_income + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  ggtitle("Monthly Energy Cost vs. Annual Income") + 
  plot_spacer() + 
  contour_plot + theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  mean_energy_cost + coord_flip() + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
```




```{r echo=FALSE,results='hide',fig.keep='all'}
predictor_vars <- c(
#"county_id",
"occupancy_type",
"primary_heating_fuel",
#"number_of_units",
#"year_constructed",
"min_age",
"min_units",
"detached",
"pct_non_male",
"median_age",
"pct_no_high_school",
"pct_non_white",
"median_income",
#"utility_id",
#"percent_tract_served",
"ownership_type",
#"balancing_authority",
#"retail_revenue_thousands",
#"retail_sales_MWh",
#"retail_customers",
"implied_price_kWh",
"revenue_per_customer",
"mean_energy_burden"
)

household_characteristics <- c(
  "occupancy_type",
  "primary_heating_fuel",
  #"number_of_units",
  #"year_constructed",
  "min_age",
  "min_units",
  "detached",
  "ownership_type",
  "implied_price_kWh",
  "mean_energy_burden"
)

pair_plot <- ggpairs(data %>% drop_na(), aes(colour = ownership_type, weight=households, alpha = 0.4),
                     columns = household_characteristics,
                     upper = list(continuous = wrap("cor", size = 2)),
                     progress=FALSE)

pair_plot
```


```{r}
ggsave("./household_pair_plot.png", plot=pair_plot, width=10,height=7.5,units="in")
```


```{r}
demographic_characteristics <- c(
  "pct_non_male",
  "median_age",
  "pct_no_high_school",
  "pct_non_white",
  "median_income",
  "ownership_type",
  "implied_price_kWh",
  "mean_energy_burden"
)

pair_plot <- ggpairs(data %>% drop_na(), aes(colour = ownership_type, alpha = 0.4),
                     columns = demographic_characteristics,
                     upper = list(continuous = wrap("cor", size = 2)),
                     progress=FALSE)

pair_plot
```

```{r}
ggsave("./demographic_pair_plot.png", plot=pair_plot, width=10,height=7.5,units="in")
```

```{r}
library(tigris)
options(tigris_use_cache = FALSE)

census_tracts <- tracts(cb = TRUE, year = 2016, class="sf", state="NC", refresh=TRUE)
#census_tracts <- st_transform(census_tracts, crs=st_crs(data_shp))
```

```{r}
#geo_data <- left_join(census_tracts, data, by=c("TRACTCE" = "tract_id"))
```


```{r}
geo_data %>% 
  ggplot() +
  geom_sf(aes(fill=ownership_type),size=.1) + 
  #geom_sf(data = bgs, fill = NA, color = "blue", size=.1) + 
  #geom_sf(data = census_tracts, fill = NA, color = "red", size=.2) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "top")
```


## Does `year_constructed` or `min_age` better explain the energy burden variance?

### Building Age vs Energy Burden

```{r}
summary(lm(formula=mean_energy_burden ~ year_constructed, data = data))
```

```{r}
summary(lm(formula=mean_energy_burden ~ min_age, data = data))
```

### Building Age vs Energy Cost

```{r}
summary(lm(formula=mean_energy_cost ~ year_constructed, data = data))
```

```{r}
summary(lm(formula=mean_energy_cost ~ min_age, data = data))
```

### Building Age vs Energy Poverty

```{r}
summary(glm(energy_poverty ~ year_constructed, data = data, family = binomial('logit')))
```

```{r}
summary(glm(energy_poverty ~ min_age, data = data, family = binomial('logit')))
```

## Does `number_of_units` or `min_units` better explain the energy burden variance?

### Building Units vs Energy Burden

```{r}
summary(lm(formula=mean_energy_burden ~ number_of_units, data = data))
```

```{r}
summary(lm(formula=mean_energy_burden ~ min_units + detached, data = data))
```

### Building Units vs Energy Cost

```{r}
summary(lm(formula=mean_energy_cost ~ number_of_units, data = data))
```

```{r}
summary(lm(formula=mean_energy_cost ~ min_units + detached, data = data))
```

### Building Units vs Energy Poverty

```{r}
summary(glm(energy_poverty ~ number_of_units, data = data, family = binomial('logit')))
```

```{r}
summary(glm(energy_poverty ~ min_units + detached, data = data, family = binomial('logit')))
```

# Simple Models

First, I decide on which variable are of interest and which predictors to use.

```{r}
response_var <- "energy_poverty"
model_type <- "classification" #regression

predictor_vars <- c(
#"county_id",
"occupancy_type",
"primary_heating_fuel",
"number_of_units",
"year_constructed",
"min_age",
"min_units",
"detached",
"pct_non_male",
"median_age",
"pct_no_high_school",
"pct_non_white",
"median_income",
#"utility_id",
"percent_tract_served",
"ownership_type",
"balancing_authority",
"retail_revenue_thousands",
"retail_sales_MWh",
"retail_customers",
"implied_price_kWh",
"revenue_per_customer"
)

# predictor_vars <- c("min_age", "number_of_units", "primary_heating_fuel", "occupancy_type")
```


```{r}
model_data <- clean_data[,c(response_var, predictor_vars)]

model_data[,response_var] <- factor(model_data[[c(response_var)]], levels=unique(model_data[[c(response_var)]]), ordered = FALSE)

model_formula <- as.formula(paste0(response_var," ~ ",paste(predictor_vars, collapse=" + ")))
```


Then, I split the data into training and testing sets.

```{r}
set.seed(13)
training_size <- round(.75 * nrow(model_data))  # training set size
indices = sample(1:nrow(model_data), training_size)

data_train <- clean_data[indices,]
data_test <- clean_data[-(indices),]

training_set <- model_data[indices,]
testing_set <- model_data[-(indices),]

dim(data_train)

#prop.table(table(data_train$primary_heating_fuel))

prop.table(table(data_train$energy_poverty))
prop.table(table(data_test$energy_poverty))
```

Then I define the classification algorithms to try.

```{r}
# Always False
# Logistic Regression
# Linear Discriminant Analysis
# Principle Component Analysis
# K-Nearest Neighbors Classification
# Support Vector Machines
# Standard Decision Trees
# Random Forest
```

Then I train and test each of them.

```{r}
test_accuracy <- function(){
  cm <- confusionMatrix(predictions, training_set[[response_var]])
}
```


## Always False

```{r}
false_in_sample_predictions <- factor(rep(FALSE, nrow(training_set)), levels=c("TRUE","FALSE"), ordered=FALSE)

cm <- confusionMatrix(false_in_sample_predictions, training_set[[response_var]])
```


```{r}
false_out_of_sample_predictions <- rep(FALSE, nrow(testing_set))

false_out_of_sample_performance <- sum(false_out_of_sample_predictions == testing_set$energy_poverty)/nrow(testing_set)
print(paste0("The false model's out-of-sample accuracy is ",label_percent()(false_out_of_sample_performance)))
```

## Logistic Regression

```{r}
logit_model <- glm(model_formula, data = training_set, family = binomial('logit'))
summary(logit_model)
```

```{r}
glm_in_sample_predictions <- predict.glm(object = logit_model,
                           newdata = training_set,
                           type = "response") > .5

glm_in_sample_performance <- sum(glm_in_sample_predictions == training_set$energy_poverty)/nrow(training_set)
print(paste0("The logistic model's in-sample performance is ",label_percent()(glm_in_sample_performance)))
```

```{r}
glm_out_of_sample_predictions <- predict.glm(object = logit_model,
                           newdata = testing_set,
                           type = "response") > .5

glm_out_of_sample_performance <- sum(glm_out_of_sample_predictions == testing_set$energy_poverty)/nrow(testing_set)
print(paste0("The logistic model's out-of-sample performance is ",label_percent()(glm_out_of_sample_performance)))
```


## Linear Discriminant Analysis

```{r}
lda_model <- lda(energy_poverty ~ ., data = training_set)
summary(lda_model)
```

```{r}
lda_in_sample_predictions <- predict(lda_model, newdata=training_set)

lda_in_sample_performance <- sum(lda_in_sample_predictions$class == training_set$energy_poverty)/nrow(training_set)
print(paste0("The LDA model's in-sample performance is ",label_percent()(lda_in_sample_performance)))


```

```{r}
lda_out_of_sample_predictions <- predict(lda_model, newdata=testing_set)

lda_out_of_sample_performance <- sum(lda_out_of_sample_predictions$class == testing_set$energy_poverty)/nrow(testing_set)
print(paste0("The LDA model's out-of-sample performance is ",label_percent()(lda_out_of_sample_performance)))
```


## K-Nearest Neighbors Classification

```{r}
knn(train_data_observations, train_data_observations, cl = train_data_classifiers, k=3)
```


## Support Vector Machines

```{r}

```


# Decision Trees

## Standard Decision Trees

```{r}
library(rpart)
library(rpart.plot)
library(neuralnet)

fit = rpart(energy_poverty~., data = training_set, method = 'class')
rpart.plot(fit, extra = 106)
summary(fit)

# Error rate
sum(predict(fit, testing_set, type='class') != testing_set$energy_poverty)/nrow(testing_set)
```


```{r}
fit = rpart(primary_heating_fuel~., data = training_set, method = 'class')
rpart.plot(fit, box.palette = "blue")#, extra = 106)
summary(fit)
```

## Standard Decision Trees for Regression

```{r}
reg_subset_predictors = c(predictor_vars, "mean_energy_burden")
reg_subset_predictors = reg_subset_predictors[!predictor_vars %in% c("energy_poverty", "number_of_units",
"year_constructed")]

reg_data_train_subset = data_train[, reg_subset_predictors]
reg_data_test_subset = data_test[, reg_subset_predictors]
tree = rpart(mean_energy_burden~., data = reg_data_train_subset, method = 'anova')
rpart.plot(tree, extra = 100)
summary(tree)
```

```{r}
RMSE = mean((reg_data_test_subset$mean_energy_burden - predict(tree, reg_data_test_subset))^2) %>% sqrt()
RMSE
```


## Random Forest

```{r}
library(randomForest)
rf_model <- randomForest(
  formula = as.factor(energy_poverty) ~ .,
  data=training_set#,
  #na.action=na.exclude
)
summary(rf_model)
```


```{r}
rf_in_sample_predictions <- predict(rf_model, newdata=training_set)

rf_in_sample_performance <- sum(rf_in_sample_predictions == training_set$energy_poverty)/nrow(training_set)
print(paste0("The Random Forest model's in-sample performance is ",label_percent()(rf_in_sample_performance)))
```

```{r}
rf_out_of_sample_predictions <- predict(rf_model, newdata=testing_set)

rf_out_of_sample_performance <- sum(rf_out_of_sample_predictions == testing_set$energy_poverty)/nrow(testing_set)
print(paste0("The Random Forest model's out-of-sample performance is ",label_percent()(rf_out_of_sample_performance)))
```

## Other Simple Models

```{r}
summary(lm(formula = mean_energy_cost ~ annual_income,
   data=data))
```

```{r}
summary(lm(formula= mean_energy_cost ~ income_bracket +
             occupancy_type + 
             primary_heating_fuel + 
             number_of_units + 
             year_constructed,
   data=data))
```

```{r}
summary(lm(formula= mean_energy_cost ~ annual_income + 
             min_units + 
             occupancy_type + 
             detached + 
             primary_heating_fuel + 
             min_age,
   data=data))
```

```{r}
summary(lm(formula= mean_energy_burden ~ annual_income + 
             min_units + 
             occupancy_type + 
             detached + 
             primary_heating_fuel + 
             min_age,
   data=data))
```

```{r}
summary(lm(formula= mean_energy_burden ~ income_bracket +
             occupancy_type + 
             primary_heating_fuel + 
             number_of_units + 
             year_constructed,
   data=data))
```


# EDA techniques to integrate

+ what is proportion of spending on by fuel type

2. Print the top 10 AQI values and their corresponding sites and dates

```{r}
data %>% 
  arrange(desc(mean_energy_burden)) %>% 
  select(mean_energy_burden, 
         annual_income, 
         mean_energy_cost, 
         income_bracket, 
         households,
         number_of_units) %>% 
  slice(1:10)
```

```{r}
data %>% 
  select(mean_energy_burden, 
         annual_income, 
         mean_energy_cost, 
         income_bracket, 
         households,
         number_of_units) %>% 
```


```{r}
data %>%
  arrange(desc(mean_energy_cost)) %>%
  select(mean_energy_cost, annual_income, mean_energy_burden) %>%
  slice(1:20)
```

```{r}
ggplot(data=data,
       aes(x=mean_energy_burden,y=households)) + 
      geom_bar(stat="identity")
```


create a unqiue id for each site identifier

```{r eval=FALSE}
data <- mutate(data, cohort_id = paste(geo_id, , sep="_")) 
head(ozone_airq_df2$SiteID)
```

### Group_by

```{r}
group_by(ozone_airq_df2, SiteID) %>%
  filter(n() < 30)
```

```{r}
group_by(ozone_airq_df2, stcofips, dateL) %>%
  summarize(maxAQI=max(AQI))
```

```{r}
county_summary_df <- group_by(ozone_airq_df2, stcofips, dateL) %>%
  summarize(maxAQI=max(AQI),
            State_Name=first(State_Name), # We can do this because we know that there is only one state corresponding to a stcofips, i.e. counties are embedded in the state.
            County_Name=first(County_Name), # We also need these variables to carry over to the next step.
            CBSA_Name=first(CBSA_Name)) %>%
  group_by(stcofips) %>%
  summarize(AQIgt100 = sum(maxAQI>=100), 
            numDays= n(), 
            percAQIgt100 = AQIgt100/numDays, 
            State_Name=first(State_Name), 
            County_Name=first(County_Name), 
            CBSA_Name=first(CBSA_Name)
            )

county_summary_df
```

## Basic Plotting

```{r}
g1 <- ozone_airq_df2 %>% 
  filter(stcofips == "06059") %>% # Orange County
  ggplot() + 
  geom_point(aes(x=dateL, y=AQI, color=SiteID)) +
  geom_smooth(aes(x=dateL, y=AQI, color=SiteID), method="loess")+
  scale_colour_brewer(palette = "Set2") + 
  labs(x = "Month", y = "Air Quality Index")

library(plotly)
ggplotly(g1)
```

```{r}
ozone_airq_df2 %>% 
  filter(stcofips == "06059") %>% # Orange County
  ggplot() + 
  coord_polar(theta = "x")+
  geom_point(aes(x=dateL, y=AQI, color=SiteID), alpha=.5, show.legend = FALSE) +
  geom_smooth(aes(x=dateL, y=AQI, color=SiteID), se=FALSE)+
  scale_colour_brewer(palette = "Dark2") + 
  labs(x = "Month", y = "Air Quality Index")
```

## Faceting

```{r}
g1 <- ozone_airq_df2 %>% 
  filter(State_Name == "California") %>% 
  ggplot() + 
  geom_smooth(aes(x=dateL, y=AQI, color=SiteID), method="loess",  se=FALSE)+
  scale_colour_grey(end = 0)+
  facet_wrap(~stcofips)+
  labs(x = "Month", y = "Air Quality Index") + 
  theme_bw() + 
  theme(axis.text.x=element_blank(),
        legend.position="none")

 ggplotly(g1)
```

### Exercise

Facet CA counties by CBSA (i.e CBSA in rows and Counties in Columns) and produce the same graph as above.

```{r}
g1 <- ozone_airq_df2 %>% 
  filter(State_Name == "California") %>% 
  ggplot() + 
  geom_smooth(aes(x=dateL, y=AQI, color=SiteID), method="loess",  se=FALSE)+
  scale_colour_grey(end = 0)+
  facet_wrap(vars(CBSA_Name,County_Name))+
  #facet_grid(rows = vars(CBSA_Name), cols = vars(County_Name))+
  labs(x = "Month", y = "Air Quality Index") + 
  theme_bw() + 
  theme(axis.text.x=element_blank(),
        legend.position="none")

 ggplotly(g1)
```


Using facetting as above, limit the x-axis to summer months.

Thanks to (stackoverflow)[https://stackoverflow.com/a/14162999]

```{r}
g1 <- ozone_airq_df2 %>% 
  filter(State_Name == "California") %>% 
  ggplot() + 
  geom_smooth(aes(x=dateL, y=AQI, color=SiteID), method="loess",  se=FALSE)+
  xlim(as.Date(c('2017-06-21', '2017-09-21'), format="%Y-%m-%d") )+
  scale_colour_grey(end = 0)+
  facet_wrap(~stcofips)+
  labs(x = "Month", y = "Air Quality Index") + 
  theme_bw() + 
  theme(axis.text.x=element_blank(),
        legend.position="none")

 ggplotly(g1)
```

```{r}
g1 <- ozone_airq_df2 %>% 
  filter(State_Name == "California")  %>% 
  ggplot() + 
  geom_smooth(aes(x=dateL, y=AQI, group=SiteID, color=stcofips), method="loess", se=FALSE)+
  labs(x = "Month", y = "Air Quality Index") + 
  theme_bw() 

ggplotly(g1)
```

## Statistical Summaries By Group

```{r}
g1 <- ozone_airq_df2 %>% 
  filter(stcofips == "06037") %>% 
  ggplot() + 
  geom_point(aes(x=Site_Num, y=AQI), show.legend = F)+
  stat_summary(aes(x=Site_Num, y=AQI), fun.y='median', colour = "red", size = 2)+
  stat_summary(aes(x=Site_Num, y=AQI),fun.y='mean', colour = "green", size = 3, shape=3)+
  labs(x = "Site", y = "Air Quality Index", title="Los Angeles County") + 
  theme_bw() 

ggplotly(g1)
```

```{r}
g1 <- ozone_airq_df2 %>% 
  filter(stcofips == "06037") %>% 
  ggplot() + 
  geom_violin(aes(x=Site_Num, y=AQI), show.legend = F)+
  stat_summary(aes(x=Site_Num, y=AQI), fun.y='median', colour = "red", size = 2)+
  stat_summary(aes(x=Site_Num, y=AQI),fun.y='mean', size = 3, shape=3)+
  labs(x = "Site", y = "Air Quality Index", title="Los Angeles County") + 
  theme_bw() 

ggplotly(g1)
```

## Cloropleth Map

Needed to install rgdal from source


```{r}
library(tigris)
library(rgdal)

ctys_spdf <- counties(cb=TRUE) #Only generalised boundaries are requiredl
```

```{r}
names(ctys_spdf@data)
# [1] "STATEFP"  "COUNTYFP" "COUNTYNS" "AFFGEOID" "GEOID"    "NAME"     "LSAD"    
# [8] "ALAND"    "AWATER"
nrow(ctys_spdf@data)
# [1] 3233
proj4string(ctys_spdf)
# [1] "+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs"
ctys_spdf <- merge(ctys_spdf, county_summary_df, by.x="GEOID", by.y='stcofips', all.x=FALSE) #Make sure to merge with spdf, not spdf@data. Otherwise, the order is jumbled.

nrow(ctys_spdf@data)
# [1] 785

# Because the presence of 0's create a problem for quantiles (some of the quantiles are the same) let's separate them out while calculating breaks.

Qpal <- colorQuantile(
  palette = "Reds", n = 5,
  domain = ctys_spdf$percAQIgt100[ctys_spdf$percAQIgt100>0]
)


labels <- sprintf(
  "County: %s <br/> AQI>100 days: <strong>%s</strong> %%",
  paste(ctys_spdf$County_Name, ctys_spdf$State_Name, sep=","),prettyNum(ctys_spdf$percAQIgt100*100, digits=2)
) %>% lapply(htmltools::HTML)

m <-  leaflet(ctys_spdf) %>%
  addProviderTiles(providers$Stamen.TonerLines, group = "Basemap") %>%
   addProviderTiles(providers$Stamen.TonerLite, group = "Basemap") %>%
       addPolygons(color = "#CBC7C6", weight = 1, smoothFactor = 0.5,
              opacity = 1.0, fillOpacity = 0.5,
             fillColor = Qpal(ctys_spdf$percAQIgt100),
              highlightOptions = highlightOptions(color = "green", weight = 2, bringToFront = TRUE),
             label = labels,
             labelOptions = labelOptions(
               style = list("font-weight" = "normal", padding = "3px 8px"),
               textsize = "15px",
               direction = "auto"),
             group = "Counties"
             )%>%
  addLayersControl(
    overlayGroups = c("Counties", 'Basemap'),
    options = layersControlOptions(collapsed = FALSE)
      ) 

  
  m 
```