---
title: 'Data Munging: Electric Retail Service Territories'
author: "Eric Scheier"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(rgdal)
library(maps)
library(tigris)
library(tidycensus)
library(readxl)

options(tigris_use_cache = FALSE)

```
# Introduction

# Acquiring Data

```{r}
# https://hifld-geoplatform.opendata.arcgis.com/datasets/c4fd0b01c2544a2f83440dab292f0980_0
erst_name <- "Electric_Retail_Service_Territories"

url_list <- list(
  "csv" = "https://opendata.arcgis.com/datasets/c4fd0b01c2544a2f83440dab292f0980_0.csv?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D",
  "kml" = "https://opendata.arcgis.com/datasets/c4fd0b01c2544a2f83440dab292f0980_0.kml?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D",
  "zip" = "https://opendata.arcgis.com/datasets/c4fd0b01c2544a2f83440dab292f0980_0.zip?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D",
  "gdb" = "https://opendata.arcgis.com/datasets/c4fd0b01c2544a2f83440dab292f0980_0.gdb?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D")

for (i in 1:length(url_list)){
  file_name <- names(url_list)[i]
  data_url <- url_list[[file_name]]
  desired_name <- paste0(erst_name,".",file_name)
  desired_path <- file.path(desired_name)
  
  if (!file.exists(desired_path)){
    print(paste0("Downloading ",desired_name))
    download.file(data_url,desired_path)
    if (file_name=="zip"){
      unzip(desired_path)
    }
  }
}
```


```{r}
data_shp <- st_read(paste0(erst_name,".shp"))

# https://gis.stackexchange.com/questions/64654/choosing-the-correct-value-for-proj4string-for-shapefile-reading-in-r
#shp <- st_transform(shp, "+init=epsg:3857")

summary(data_shp)
```

```{r}
str(data_shp)
```

```{r}
#plot(data_shp$geometry)
```


```{r}
data <- data_shp %>% st_drop_geometry()

str(data)
```


# Data Munging

Munging Steps:

+ Limit to North Carolina Utilities
+ Replace `-999999` with `NA`

## Limit to NC Utilities

For now I will limit by the state name of the serving utility. Soon, I will select by those territories which overlap with the NC state boundary.

```{r}
#plot(data_shp$geometry[data$STATE=="NC" & data$TYPE=="INVESTOR OWNED"])
```

## Reclassify `STATE` and `MUNICIPAL MARKETING AUTHORITIES` as `MUNICUPAL` service providers

```{r}
#plot(counties$geometry)
#counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))
#counties <- subset(counties, grepl("north carolina", counties$ID))


#plot(states$geometry)
#states <- st_as_sf(map("state", plot=FALSE, fill=TRUE))
#states <- subset(states, grepl("north carolina", states$ID))

#block_groups
bgs <- block_groups(cb = TRUE, year = 2016, class="sf", state="NC", refresh=TRUE)
bgs <- st_transform(bgs, crs=st_crs(data_shp))
#plot(bgs$geometry)

mean_bg_area <- st_area(bgs) %>% mean()

#tracts
census_tracts <- tracts(cb = TRUE, year = 2016, class="sf", state="NC", refresh=TRUE)
census_tracts <- st_transform(census_tracts, crs=st_crs(data_shp))
#plot(census_tracts$geometry)

#zctas

#counties
counties <- counties(state = 'NC', cb = FALSE, year=2016, resolution = "500k", class="sf", refresh=TRUE)
counties <- st_transform(counties, crs=st_crs(data_shp))

#state
states <- states(cb = FALSE, resolution = "500k", year = 2016, class="sf", refresh=TRUE)
states <- subset(states, GEOID=="37")
states <- st_transform(states, crs=st_crs(data_shp))
```

```{r}
nc_shp <- st_intersection(data_shp, states)
#plot(nc_shp$geometry)
```



```{r}
nc_shp %>% filter(TYPE=="MUNICIPAL MKTG AUTHORITY") %>% 
  ggplot() + 
  geom_sf(aes(fill=SOURCE, alpha=.2), color=NA) + 
  geom_sf(data = counties, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position="bottom") + 
  guides(fill = guide_legend(ncol = 1))
```

```{r}
nc_shp %>% filter(TYPE=="STATE") %>% 
  ggplot() + 
  geom_sf(aes(fill=SOURCE, alpha=.2), color=NA) + 
  geom_sf(data = counties, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position="bottom") + 
  guides(fill = guide_legend(ncol = 1))
```

```{r}
nc_shp %>% filter(TYPE=="MUNICIPAL") %>% 
  ggplot() + 
  geom_sf(aes(fill=SOURCE, alpha=.2), color=NA) + 
  geom_sf(data = counties, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position="bottom") + 
  guides(fill = guide_legend(ncol = 1))
```


```{r}
muni_sources <- c("EIA 861, https://www.tva.gov/file_source/TVA/Site%20Content/Energy/tva_lpc_map.pdf",
             "EIA 861, Urban Areas - Cartographic Boundary Shapefiles - U.S. Census",
             "FERC 714, EIA 861, Urban Areas - Cartographic Boundary Shapefiles - U.S. Census",
             "EIA 860, EIA 861, Urban Areas - Cartographic Boundary Shapefiles - U.S. Census")
nc_shp %>% filter(TYPE=="MUNICIPAL" & SOURCE %in% muni_sources) %>% 
  ggplot() + 
  geom_sf(aes(fill=SOURCE, alpha=.2), color=NA) + 
  geom_sf(data = counties, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position="bottom") + 
  guides(fill = guide_legend(ncol = 1))
```



```{r}
nc_shp %>% filter(TYPE=="COOPERATIVE") %>% 
  ggplot() + 
  geom_sf(aes(fill=SOURCE, alpha=.2), color=NA) + 
  geom_sf(data = counties, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position="bottom") + 
  guides(fill = guide_legend(ncol = 1))
```

```{r}
not_coop_sources <- c("FERC 714, EIA 861, TIGER/Line Shapefiles - U.S. Census, https://www.electric.coop/wp-content/uploads/2018/09/South_Carolina_Coops-1024x791.png",
                      "EIA 861, TIGER/Line Shapefiles - U.S. Census, https://www.electric.coop/information-hub-electric-co-ops-hurricane-michael/")
nc_shp %>% filter(TYPE=="COOPERATIVE" & !(SOURCE %in% not_coop_sources)) %>% 
  ggplot() + 
  geom_sf(aes(fill=SOURCE, alpha=.2), color=NA) + 
  geom_sf(data = counties, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position="bottom") + 
  guides(fill = guide_legend(ncol = 1))
```


```{r}
nc_shp %>% filter(TYPE=="INVESTOR OWNED") %>% 
  ggplot() + 
  geom_sf(aes(fill=SOURCE, alpha=.2), color=NA) + 
  geom_sf(data = counties, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position="bottom") + 
  guides(fill = guide_legend(ncol = 1))
```

```{r}
nc_shp %>% filter(TYPE=="INVESTOR OWNED") %>% 
  ggplot() + 
  geom_sf(aes(fill=NAME, alpha=.2), color=NA) + 
  geom_sf(data = counties, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position="bottom") + 
  guides(fill = guide_legend(ncol = 1))
```


```{r}
# Territory types not applicable to North Carolina
## FEDERAL
## POLITICAL SUBDIVISION
## NOT AVAILABLE
## WHOLESALE POWER MARKETER
# nc_shp %>% filter(TYPE=="") %>% 
#   ggplot() + 
#   geom_sf(aes(fill=SOURCE, alpha=.2), color=NA) + 
#   geom_sf(data = counties, fill = NA, color = "black") + 
#   coord_sf(crs = 4326) + 
#   theme(legend.position="bottom") + 
#   guides(fill = guide_legend(ncol = 1))
```

Select relevant utilities based on analysis above


```{r}
nc_shp %>% filter((TYPE=="MUNICIPAL" & SOURCE %in% muni_sources) | 
                                 (TYPE=="COOPERATIVE" & !(SOURCE %in% not_coop_sources)) | 
                                 (TYPE=="INVESTOR OWNED")) %>% 
  ggplot() + 
  geom_sf(aes(fill=TYPE, alpha=.2), color=NA) + 
  #geom_sf(data = counties, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position="bottom") + 
  guides(fill = guide_legend(ncol = 1))
  
```

```{r}
nc_clean_shp <- subset(nc_shp, (TYPE=="MUNICIPAL" & SOURCE %in% muni_sources) | 
                                 (TYPE=="COOPERATIVE" & !(SOURCE %in% not_coop_sources)) | 
                                 (TYPE=="INVESTOR OWNED"))
```

```{r}
nc_clean_shp %>% 
  ggplot() + 
  geom_sf(aes(fill=NAME, alpha=.2), color=NA) + 
  geom_sf(data = states, fill = NA, color = "black") + 
  coord_sf(crs = 4326) +
  theme(legend.position = "none")
  #theme(legend.position="bottom") + 
  #guides(fill = guide_legend(ncol = 1))
```

## Flatten the service territories for each provider type

#### Function to assign overlapping areas

```{r}
nc_clean_shp %>% filter(TYPE=="INVESTOR OWNED") %>% #st_sf() %>% 
  st_buffer(dist = 0) %>% 
  ggplot() + 
  geom_sf(aes(fill=ID, alpha=.2), color=NA) + 
  geom_sf(data = states, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "none")
```

```{r}
# https://github.com/r-spatial/sf/issues/824
st_no_overlap <- function(polygons) {
  
  #independent <- b %>% st_sf %>% st_intersection %>% subset(n.overlaps<=1) %>% st_cast("POLYGON")
  #overlap <- b %>% st_sf %>% st_intersection %>% subset(n.overlaps>1) %>% st_union()
  
  centroids <- polygons %>% st_centroid

     # Voronoi tesselation
     voronoi <- 
          centroids %>% 
          st_geometry() %>%
          st_union() %>%
          st_voronoi() %>%
          st_collection_extract()
     
     # Put them back in their original order
     voronoi <-
          voronoi[unlist(st_intersects(centroids,voronoi))]

     # Keep the attributes
     result <- centroids
     
     # Intersect voronoi zones with buffer zones
     st_geometry(result) <-
          mapply(function(x,y) st_intersection(x,y),
                 #st_buffer(st_geometry(centroids),dist), 
                 polygons$geometry,
                 voronoi,
                 SIMPLIFY=FALSE) %>%
          st_sfc(crs=st_crs(centroids))

     result
}

unoverlap_territories <- function(territories, min_area){
  all_polygons <- territories %>% st_cast("MULTIPOLYGON") %>% st_cast("POLYGON") %>% 
    st_buffer(dist = 0) %>% st_set_precision(1e5) %>%  lwgeom::st_make_valid()
  
  all_polygons <- all_polygons[st_area(all_polygons)>min_area,,drop=FALSE]
  
  final_b <- st_no_overlap(all_polygons)
  
  independent <- all_polygons %>% st_sf %>% st_intersection %>% subset(n.overlaps<=1) %>% 
    st_cast("MULTIPOLYGON") %>% st_cast("POLYGON") %>% st_buffer(dist = 0) %>% 
    st_set_precision(1e5) %>%  lwgeom::st_make_valid()
  
  # combine by ID
  combined_territories <- rbind(independent %>% select(names(final_b)), final_b)
  
  # merge independent with final_b by ID
  merged_territories <- combined_territories %>% group_by(ID) %>% 
    summarize() %>% ungroup() %>% st_buffer(dist = 0) %>% 
    st_set_precision(1e5) %>% lwgeom::st_make_valid() %>% 
    st_difference() %>% st_intersection() %>% 
    subset(n.overlaps==1)
  
  nice_merged_territories <- merged_territories %>% st_buffer(dist = 0) %>% st_set_precision(1e5) %>%  lwgeom::st_make_valid()
  
  
  return(nice_merged_territories)
}

#overlap <- all_polygons %>% st_sf %>% st_intersection %>% subset(n.overlaps>1) %>% st_union()

# abort if there are no overlaps
#if(identical(as.numeric(st_area(overlap)), numeric(0))){
#  return(b)
#}
```

### IOUs

```{r}
min_area <- st_area(bgs) %>% mean #median

iou_shp <- nc_clean_shp %>% filter(TYPE=="INVESTOR OWNED") %>% st_buffer(dist = 0)

nice_ious <- unoverlap_territories(territories = iou_shp,
                                   min_area = min_area)
```


```{r}
nice_ious %>% 
  ggplot() + 
  geom_sf(aes(fill=ID, alpha=.2), color="red") + 
  #geom_sf(data = states, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "bottom")
```

### Muni's


```{r}
nc_clean_shp %>% filter(TYPE=="MUNICIPAL") %>% 
  ggplot() + 
  geom_sf(aes(fill=NAME, alpha=.2), color=NA) + 
  geom_sf(data = states, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  #theme(legend.position = "none")
  theme(legend.position="bottom")# + 
  #guides(fill = guide_legend(ncol = 1))
```

```{r}
muni_shp <- nc_clean_shp %>% filter(TYPE=="MUNICIPAL") %>% st_buffer(dist = 0)

nice_munis <- unoverlap_territories(territories = muni_shp,
                                   min_area = min_area)
```


```{r}
nice_munis %>% 
  ggplot() + 
  geom_sf(aes(fill=ID, alpha=.2), color="red") + 
  #geom_sf(data = states, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "bottom")
```


### Coops

```{r}
nc_clean_shp %>% filter(TYPE=="COOPERATIVE") %>% #st_sf() %>% 
  st_buffer(dist = 0) %>% 
  ggplot() + 
  geom_sf(aes(fill=NAME, alpha=.2), color=NA) + 
  geom_sf(data = states, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  #theme(legend.position = "none")
  theme(legend.position="bottom")# + 
  #guides(fill = guide_legend(ncol = 1))
```

```{r}
coop_shp <- nc_clean_shp %>% filter(TYPE=="COOPERATIVE") %>% st_buffer(dist = 0)

nice_coops <- unoverlap_territories(territories = coop_shp,
                                   min_area = min_area)
```


```{r}
nice_coops %>% 
  ggplot() + 
  geom_sf(aes(fill=ID, alpha=.2), color="red") + 
  #geom_sf(data = states, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "bottom")
```


### Combine All Territories Together

```{r}
combine_territories <- function(t_front, t_back){
  t_front_polygons <- t_front %>% st_cast("MULTIPOLYGON") %>% st_cast("POLYGON") %>% 
    st_buffer(dist = 0) %>% st_set_precision(1e5) %>%  lwgeom::st_make_valid()
  
  t_back_polygons <- t_back %>% st_cast("MULTIPOLYGON") %>% st_cast("POLYGON") %>% 
    st_buffer(dist = 0) %>% st_set_precision(1e5) %>%  lwgeom::st_make_valid()
  #final_b <- st_no_overlap(all_polygons)
  #final_b <- st_intersect()
  
  back_independent <- st_difference(st_union(t_back_polygons), st_union(t_front_polygons))# %>% 
    #st_cast("MULTIPOLYGON") %>% st_cast("POLYGON")
  
  t_back_independent <- st_intersection(t_back_polygons, back_independent) %>% 
    subset(n.overlaps<=1) %>% st_set_precision(1e5) %>%  lwgeom::st_make_valid() %>% st_buffer(dist = 0)
    #st_intersection(t_back_polygons, st_difference(st_union(t_back_polygons), st_union(t_front_polygons))) %>% 
    #subset(n.overlaps<=1)# %>% 
    #st_cast("MULTIPOLYGON") %>% st_cast("POLYGON") %>% 
    #st_buffer(dist = 0) %>% st_set_precision(1e5) %>%  lwgeom::st_make_valid()
  
  # combine by ID
  combined_territories <- rbind(t_back_independent %>% select(names(t_front)), t_front)
  
  # merge independent with final_b by ID
  merged_territories <- combined_territories %>% group_by(ID) %>% 
    summarize() %>% ungroup() %>% st_buffer(dist = 0) %>% 
    st_set_precision(1e5) %>% lwgeom::st_make_valid() %>% 
    st_difference() %>% st_intersection() %>% 
    subset(n.overlaps==1)
  
  nice_merged_territories <- merged_territories %>% st_buffer(dist = 0) %>% st_set_precision(1e5) %>%  lwgeom::st_make_valid()
  
  
  return(nice_merged_territories)
}

combined_muni_coops <- combine_territories(nice_munis, nice_coops)
combined_coop_ious <- combine_territories(nice_coops, nice_ious)

#full_combine <- combine_territories(combined_muni_coops, nice_ious)
full_combine <- combine_territories(nice_munis, combined_coop_ious)
```


```{r}
full_combine %>% 
  ggplot() + 
  geom_sf(aes(fill=ID, alpha=.2), color=NA) + 
  #geom_sf(data=nice_munis, aes(fill=ID, alpha=.2), color="green") + 
  #geom_sf(data = states, fill = NA, color = "black") + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "bottom")
```

```{r}
st_write(full_combine, "full_combined_ersts.geojson", delete_dsn = TRUE)
```


## Assign each census block group to a service territory

```{r}
ggplot() + 
  geom_sf(data=full_combine, aes(fill=ID, alpha=.2), color=NA) + 
  geom_sf(data = bgs, fill = NA, color = "blue", size=.1) + 
  geom_sf(data = census_tracts, fill = NA, color = "red", size=.2) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "none")
```

```{r}
block_group_utilities <- st_join(bgs, full_combine, left=TRUE, largest=TRUE)
```


```{r}
ggplot() + 
  geom_sf(data=block_group_utilities, aes(fill=ID, alpha=.2), color=NA) + 
  #geom_sf(data = bgs, fill = NA, color = "blue", size=.1) + 
  #geom_sf(data = census_tracts, fill = NA, color = "red", size=.2) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "none")
```

```{r}
st_write(block_group_utilities, "block_group_utilities.geojson", delete_dsn = TRUE)
```


```{r}
# B00002_001E	Estimate!!Total	UNWEIGHTED SAMPLE HOUSING UNITS
# B25001_001E	Estimate!!Total	HOUSING UNITS

# housing_density <- get_acs(state = "NC",
#                            geography = "block group",
#                            variables = c("B25001_001"),
#                            year = 2016,
#                            output = "wide",
#                            geometry = FALSE)

acs_variables <- list(
  total_households = sym("B25001_001")
)

acs_features <- get_acs(state = "NC", 
                    geography = "block group",
                    variables = as.character(unlist(acs_variables)),
                    year = 2016,
                    output = "tidy",
                    geometry = FALSE)

acs_features <- pivot_wider(acs_features, id_cols=c("GEOID"),
                            names_from=c("variable"),
                            #names_prefix = "acs_",
                            values_from=c("estimate"))

acs_features <- rename(acs_features, !!!acs_variables)
head(acs_features)

```


```{r}
acs_block_group_utilities <- st_sf(left_join(acs_features, block_group_utilities, by="GEOID"))

acs_block_group_utilities %>% filter(is.na(ID)) %>% summarise(covered=sum(total_households))

acs_block_group_utilities %>% filter(is.na(ID)) %>% 
  ggplot() +
  geom_sf(aes(fill="red", alpha=.2), color="blue", size=.2) + 
  geom_sf(data = bgs, fill = NA, color = "blue", size=.1) + 
  #geom_sf(data = census_tracts, fill = NA, color = "red", size=.2) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "none")
```

```{r}
tract_utilities <- acs_block_group_utilities %>% st_set_geometry(NULL) %>% mutate(ID=fct_explicit_na(ID)) %>% 
  group_by(TRACTCE, ID) %>% 
  summarise(tract_households_served = sum(total_households)) %>% 
  left_join(acs_block_group_utilities %>% st_set_geometry(NULL) %>% group_by(TRACTCE) %>% 
              summarise(tract_households = sum(total_households)), by="TRACTCE") %>% 
  mutate(percent_tract_served = tract_households_served / tract_households) %>% 
  group_by(TRACTCE) %>% slice(which.max(percent_tract_served))# %>% 
  #arrange((percent_tract_served)) %>% head(50)# slice(1:10)

```

```{r eval=FALSE}
left_join(census_tracts, tract_utilities %>% filter(ID=="(Missing)"), by="TRACTCE") %>% 
  ggplot() +
  geom_sf(aes(fill=ID, alpha=.2), color="blue", size=.2) + 
  #geom_sf(data = bgs, fill = NA, color = "blue", size=.1) + 
  #geom_sf(data = census_tracts, fill = NA, color = "red", size=.2) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "none")
```


```{r}
left_join(census_tracts, tract_utilities %>% filter(percent_tract_served<.5), by="TRACTCE") %>% 
  ggplot() +
  geom_sf(aes(fill=ID, alpha=.2), color="blue", size=.2) + 
  #geom_sf(data = bgs, fill = NA, color = "blue", size=.1) + 
  #geom_sf(data = census_tracts, fill = NA, color = "red", size=.2) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "none")
```

```{r}
# Percent of households in a tract that is served entirely by 1 utility
sum((tract_utilities$percent_tract_served==1) * tract_utilities$tract_households_served)/sum(tract_utilities$tract_households)

# Percent of households served by the utility assigned to their tract
sum(tract_utilities$tract_households_served)/sum(tract_utilities$tract_households)
```


```{r}
left_join(census_tracts, tract_utilities, by="TRACTCE") %>% 
  ggplot() +
  geom_sf(aes(fill=percent_tract_served), color="blue", size=.2) + 
  #geom_sf(data = bgs, fill = NA, color = "blue", size=.1) + 
  #geom_sf(data = census_tracts, fill = NA, color = "red", size=.2) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "top")
```

```{r}
tract_utilities <- left_join(census_tracts, tract_utilities, by="TRACTCE")
```


# Join EIA EIA-861 Data

```{r}
# Download the EIA-861 Data
zip_url <- "https://www.eia.gov/electricity/data/eia861/archive/zip/f8612016.zip"

desired_path <- file.path("Frame_2016.xlsx")

if (!file.exists(desired_path)){
  print(paste0("Downloading ",desired_name))
  temp <- tempfile()
  download.file(zip_url,temp)
  unzip(temp)
  unlink(temp)
}
```


```{r}
# Get the sheets we want
# Filter by the variables we want
#desired_fields <- list(Frame=c(Utility Number	Utility Name	Ownership Code	Ownership)
frame <- read_excel(file.path("Frame_2016.xlsx"))
sales <- read_excel(path=file.path("Sales_Ult_Cust_2016.xlsx"),
                    sheet="States",
                    skip=2)
```


```{r}
# Ownership Type
frame <- select(frame,
               `Utility Number`,
               `Utility Name`,
               `Ownership`,
               `Short Form`,
               `Sales to Ultimate Customers`)

frame <- rename(frame,
               ID=`Utility Number`,
               utility_name=`Utility Name`,
               ownership_type=`Ownership`) %>% 
  mutate(ID=as.factor(ID))

type <- select(frame, `ID`,`utility_name`, `ownership_type`)
```


```{r}
## Retail Price
## Retail Customers
## Retail Load

sales <- sales %>% select(ID=`Utility Number`,
                          utility_name=`Utility Name`,
                          ownership_type=`Ownership`,
                 part=`Part`,
                 service_type=`Service Type`,
                 imputed=`Data Type\r\nO = Observed\r\nI = Imputed`,
                 state=`State`,
                 balancing_authority=`BA_CODE`,
                 retail_revenue_thousands=`Thousand Dollars...10`,
                 retail_sales_MWh=`Megawatthours...11`,
                 retail_customers=`Count...12`) %>% 
  filter(ID!=99999, retail_revenue_thousands!=0) %>%  
  mutate(ID=as.character(ID),
                  imputed=if_else(imputed=="O",FALSE,TRUE,NA),
                  retail_revenue_thousands=as.numeric(retail_revenue_thousands),
                  retail_sales_MWh=as.numeric(retail_sales_MWh),
                  retail_customers=as.numeric(retail_customers),
                  implied_price_kWh=retail_revenue_thousands/retail_sales_MWh,
                  revenue_per_customer=retail_revenue_thousands/(1000*retail_customers))

state_sales <- sales %>% filter(state=="NC")
```

```{r}
ggplot(state_sales, aes(x=implied_price_kWh, color=ownership_type, fill=ownership_type)) + 
  geom_histogram(binwidth = 0.005)
```

```{r}
state_sales %>% filter(ID %in% unique(tract_utilities$ID)) %>%
  ggplot(aes(x=implied_price_kWh, color=ownership_type, fill=ownership_type)) + 
  geom_histogram(binwidth = 0.005)
```

```{r}
# Join to tract_utilities based on utility "ID"
tract_utilities <- left_join(tract_utilities, state_sales, by="ID")
```

```{r}
st_write(tract_utilities, "tract_utilities.geojson", delete_dsn = TRUE)
```

```{r}
tract_utilities <- st_read("tract_utilities.geojson")
```

```{r}
tract_utilities %>% 
  ggplot() +
  geom_sf(aes(fill=ownership_type),size=.1) + 
  #geom_sf(data = bgs, fill = NA, color = "blue", size=.1) + 
  geom_sf(data = counties, fill = NA, color = "black", size=.2) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "top")
```

```{r eval=FALSE}
left_join(census_tracts, tract_utilities, by="TRACTCE") %>% 
  ggplot() +
  geom_sf(aes(fill=utility_name),size=.1) + 
  #geom_sf(data = bgs, fill = NA, color = "blue", size=.1) + 
  #geom_sf(data = census_tracts, fill = NA, color = "red", size=.2) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "right")
```

```{r}
summary(tract_utilities)
```


```{r}
# Convert Mising IDs back to NA
# ID as factor
# TRACTCE as factor (or even to geoID)
# utility_type as factor

#undo ID=fct_explicit_na(ID)
tract_utilities$ID <- ifelse(!is.na(tract_utilities$ID) & tract_utilities$ID == "(Missing)", NA, tract_utilities$ID)

final_data <- tract_utilities %>% select(tract_id=`TRACTCE`,
                                         utility_id=`ID`,
                                         `tract_households_served`,
                                         `percent_tract_served`,
                                         `utility_name`,
                                         `ownership_type`,
                                         `balancing_authority`,
                                         `retail_revenue_thousands`,
                                         `retail_sales_MWh`,
                                         `retail_customers`,
                                         `implied_price_kWh`,
                                         `revenue_per_customer`) %>% ungroup(tract_id) %>% 
  mutate(tract_id=as.factor(tract_id),
         utility_id=as.factor(utility_id),
         ownership_type=as.factor(ownership_type))


st_write(final_data, "clean_erst.geojson", delete_dsn = TRUE)

final_data_no_geometry <- st_set_geometry(final_data, NULL)
```


```{r}
# Save to csv
write_csv(final_data_no_geometry, "clean_erst.csv")
```


```{r}
data_reloaded <- read_csv(file.path(getwd(),"clean_erst.csv"),
                          col_types=cols(
                            tract_id = readr::col_factor(),
                            utility_id = readr::col_factor(),
                            percent_tract_served = readr::col_double(),
                            tract_households_served = readr::col_double(),
                            utility_name = readr::col_character(),
                            ownership_type = readr::col_factor(NULL),
                            balancing_authority = readr::col_factor(NULL),
                            retail_revenue_thousands = readr::col_double(),
                            retail_sales_MWh = readr::col_double(),
                            retail_customers = readr::col_double(),
                            implied_price_kWh = readr::col_double(),
                            revenue_per_customer = readr::col_double()))

all(data_reloaded == final_data_no_geometry, na.rm=TRUE)
```
