---
title: 'Net Household Income Earned for Every Dollar Spent on Energy'
author: "Eric Scheier"
date: "`r format(Sys.time(), '%Y-%B-%d')`"
output:
  html_document: default
  pdf_document: default
  tufte::tufte_handout: default
  word_document: default
---

```{r setup, include=FALSE}
is_final=FALSE
is_preview=TRUE
is_draft=TRUE
set.seed(123)

knitr::opts_chunk$set(comment='##', 
                      collapse=ifelse(is_preview,TRUE,!is_draft),
                      echo=ifelse(is_preview,FALSE,is_draft),
                      eval=TRUE,
                      warning=ifelse(is_preview,FALSE,is_draft),
                      error=ifelse(is_preview,FALSE,is_draft),
                      results=ifelse(is_final,'hide',ifelse(is_preview,'hide','asis')),
                      fig.keep='all',
                      message=ifelse(is_preview,FALSE,is_draft),
                      include=ifelse(is_preview,TRUE,is_draft),
                      tidy=TRUE,
                      cache=FALSE,
                      fig.margin=FALSE,
                      fig.fullwidth = FALSE
                      )
```

```{r include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
kable(knitr::opts_chunk$get() %>% enframe())
```


```{r}
library(sf)
library(MASS)
library(scales)
library(caret)
library(tidyverse)
library(spatstat)
library(GGally)
library(ggpubr)
library(patchwork)
library(mgcv)
library(tidyselect)
library(RColorBrewer)
library(wesanderson)
library(tigris)
library(readxl)
source("lead_munging.R")
```

```{r}
states <- "nc" #c("nc","sc") #  
income_metric <- "fpl15" #"ami68"
geographic_scope <- "tract" #statecitycounty

col_types <- cols(geo_id = readr::col_character(), 
                  state_id = readr::col_character(),
                  #state = readr::col_character(),
                  county_id = readr::col_character(),
                  #county = readr::col_character(),
                  tract_id = readr::col_character(),
                  puma10_code = readr::col_character(),
                  fmr_code = readr::col_character(),
                  occupancy_type = readr::col_factor(),
                  income_bracket = readr::col_factor(),
                  primary_heating_fuel = readr::col_factor(),
                  number_of_units = readr::col_factor(),
                  year_constructed = readr::col_factor(),
                  households = col_double(),
                  annual_income = col_double(),
                  electricity_spend = col_double(),
                  gas_spend = col_double(),
                  other_spend = col_double(),
                  acs_responses = col_double(),
                  min_age = col_double(),
                  min_units = col_double(),
                  detached = readr::col_factor(NULL),
                  mean_energy_cost = col_double(),
                  pct_non_male = col_double(), 
                  median_age = col_double(),
                  pct_no_high_school = col_double(),
                  pct_non_white = col_double(),
                  median_income = col_double(),
                  acs_total_households = col_double())

data <- get_multiple_states(states=states,
                            income_metric=income_metric,
                            geographic_scope=geographic_scope,
                            refresh=TRUE,
                            col_types=col_types,
                            raw=FALSE)

save_file_name <- toupper(paste(income_metric,
                                geographic_scope,
                                "sh",
                                paste(sort(states),  collapse = "_"), 
                                sep = "_"))

# clean_file_name <- paste0("clean_lead_",save_file_name,".csv")

# data <- read_csv(file.path(getwd(),clean_file_name),col_types=col_types)
```


```{r}
data$year_constructed <- fct_reorder(data$year_constructed, data$min_age)
data$number_of_units <- fct_reorder(data$number_of_units, data$min_units)
data$income_bracket <- factor(data$income_bracket, levels=sort(levels(data$income_bracket)))

data$state_name <- factor(data$state_name, levels=sort(unique(data$state_name)))
data$area_name <- factor(data$area_name, levels=sort(unique(data$area_name)))
data$state_code_usps <- factor(data$state_code_usps, levels=sort(unique(data$state_code_usps)))


# for (col_to_order in cols_to_order){
#   data[[col_to_order]] <- factor(data[[col_to_order]], 
#                                  levels=levels(data[[col_to_order]])[
#                                    sort(as.numeric(str_extract(levels(data[[col_to_order]]), "[0-9]+")), 
#                                         index.return=TRUE, 
#                                         decreasing = TRUE)[["ix"]]
#                                    ])
#   print(levels(data[[col_to_order]]))
# }
```

```{r}
#str(data) #print(kable(var_desc) %>% kable_styling(c("striped", "bordered")))
#summary(data)
```

# Calculating Additional Metrics

## Poverty

```{r echo=TRUE}
data$in_poverty <- as.factor(ifelse(data$income_bracket=="0-100%","Below FPL", "Above FPL"))
```

How many households are below the poverty line in this data?

```{r}
total_households <- sum(data$households)

number_in_poverty <- sum((data$income_bracket=="0-100%") * (data$households))

pct_in_poverty <- number_in_poverty / total_households

number_above_poverty <- sum((data$income_bracket!="0-100%") * (data$households))

pct_above_poverty <- number_above_poverty / total_households
```


## Annual Energy Spending

```{r echo=TRUE}
data$annual_energy_cost <- 12*data$mean_energy_cost
```

## Net Income

```{r echo=TRUE}
data$net_annual_income <- data$annual_income - (data$annual_energy_cost)
```

# Adding Net Energy Analysis Metrics

Most/All Net Energy Analysis metrics are composed of the gross amount of resource extracted, and the amount of resource spent in the extraction process. For households extracting income from the economy, these ratios will be composed of:

`g` = gross income
`s` = spending on energy

How are these metrics related? We are specifically curious about the relationship among households considered to be in poverty by the federal government.

```{r}
ggplot(data=data, aes(x=annual_income, y=annual_energy_cost, color=in_poverty)) + 
  geom_point(aes(size=households), alpha=0.05, show.legend = FALSE) + 
  geom_smooth(method=lm,
                se=TRUE,
                fullrange=TRUE)# + 
  #xlim(0,500000)
  
```

From these metrics we can create all of the relevant energy ratios:

## Traditional Energy Burden Indicator

`energy_burden = s/g`


```{r}
# Create the Energy Burden Indicator `mean_energy_burden`.

energy_burden_func <- function(g,s){
  s/g
}

# data$energy_burden <- data$mean_energy_cost / (data$annual_income/12.0)
data$energy_burden <- energy_burden_func(g=data$annual_income,
                                         s=12*data$mean_energy_cost)
```

Energy poverty is commonly defined as an expenditure of greater than 10% of household income on energy^[D. J. Bednar and T. G. Reames, “Recognition of and response to energy poverty in the United States,” Nature Energy, 2020.]. This level will be referred to as the "Energy Poverty Line" and translated into its relative level for the other ratios below.

```{r}
energy_burden_poverty_line <- 0.10

#For further analysis, I will add a designation of whether a cohort is, on average, in energy poverty depending on whether the mean energy burden is above `r label_percent()(energy_burden_poverty_line)`.

data$energy_burden_poverty <- as.logical(data$energy_burden > energy_burden_poverty_line)
```

## Energy Return on Investment

`eroi = g/s`

```{r}
# Create the Energy Return on Investment Indicator `eroi`
eroi_func <- function(g,s){
  g/s
}

# data$eroi <- data$annual_income / (12*data$mean_energy_cost)
data$eroi <- eroi_func(g=data$annual_income,
                       s=12*data$mean_energy_cost)

eroi_poverty_line <- eroi_func(g=1,
                               s=energy_burden_poverty_line)
```


## Net Energy Ratio (or Net Energy Return)

`ner = (g-s)/s`

```{r}
ner_func <- function(g,s){
  (g-s)/s
}

# data$ner <- (data$annual_income - (12*data$mean_energy_cost)) / (12*data$mean_energy_cost)
data$ner <- ner_func(g=data$annual_income,
                     s=12*data$mean_energy_cost)

ner_poverty_line <- ner_func(g=1,
                             s=energy_burden_poverty_line)
```

## Discretionary Energy Availability Rate

`dear = (g-s)/g`

This is equal to `1 - energy_burden`.

```{r}
dear_func <- function(g,s){
  (g-s)/g
}

# data$dear <- (data$annual_income - (12*data$mean_energy_cost)) / data$annual_income
data$dear <- dear_func(g=data$annual_income,
                       s=12*data$mean_energy_cost)

dear_poverty_line <- dear_func(g=1,
                               s=energy_burden_poverty_line)
```


# Analysis

Of the above metrics, the Net Energy Ratio (NER) and Discretionary Energy Availability Rate (DEAR) are the most desirable, The DEAR is nicely bound between 0-100%, while the NER is the standard metric of Net Energy Analysis.

EROI is a precursor of NER, while Energy Burden is the metric du jour in the energy insecurity literature.

```{r}
clean_data <- data #drop_na(data)
```

```{r eval=FALSE}
ggplot(data=clean_data, aes(x=eroi, y=dear, fill=income_bracket, color=income_bracket)) + 
  xlim(c(0,100)) + 
  ylim(c(0,1)) + 
  geom_point(aes(size=households), alpha=0.1, position="jitter") + 
  geom_hline(yintercept = .9) + 
  geom_vline(xintercept = 10) + 
  geom_smooth(method="lm",
              se=FALSE,
              fullrange=TRUE)
```

DEAR and NER are often compared to portray a phenomenon known as the "net energy cliff" Do we see the energy cliff in these data?

```{r}
ggplot(data=clean_data, aes(x=ner, y=dear, fill=in_poverty, color=in_poverty)) + 
  xlim(c(0,100)) + 
  ylim(c(0,1)) + 
  geom_point(aes(size=households), alpha=0.1, position="jitter", show.legend = FALSE) + 
  geom_hline(yintercept = .9) + 
  geom_vline(xintercept = 10) + 
  geom_smooth(method="lm",
              se=FALSE,
              fullrange=TRUE)
```


```{r}
#chart_title <- "Household Economic Return on Energy Spending"
chart_title <- "Net Energy Income Ratio"
chart_subtitle <- "Net Household Income Earned for Every Dollar Spent on Energy"

group_columns <- NULL#"income_bracket")#in_poverty
                   #"primary_heating_fuel")

metric_name <- "ner" #"energy_burden" #"ner" #"dear" #"eroi"
metric_label <- "Net Energy Income Ratio"
metric_cutoff_level <- ner_poverty_line
metric_cutoff_label <- "Energy Poverty Line"

upper_quantile_view <- .99
lower_quantile_view <- .01
```

```{r}
stat_ewcdf <- function(mapping = NULL, data = NULL,
                      geom = "step", position = "identity",
                      weight =  NULL, 
                      ...,
                      n = NULL,
                      pad = TRUE,
                      na.rm = FALSE,
                      show.legend = NA,
                      inherit.aes = TRUE) {
  layer(
    data = data,
    mapping = mapping,
    stat = StatEcdf,
    geom = geom,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      n = n,
      pad = pad,
      na.rm = na.rm,
      weight = weight,
      ...
    )
  )
}


#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
#' 

StatEcdf <- ggproto("StatEcdf", Stat,
                    compute_group = function(data, scales, weight, n = NULL, pad = TRUE) {
                      # If n is NULL, use raw values; otherwise interpolate
                      if (is.null(n)) {
                        x <- unique(data$x)
                      } else {
                        x <- seq(min(data$x), max(data$x), length.out = n)
                      }
                      
                      if (pad) {
                        x <- c(-Inf, x, Inf)
                      }
                      y <- ewcdf(data$x, weights=data$weight/sum(data$weight))(x)
                      
                      data.frame(x = x, y = y)
                    },
                    
                    default_aes = aes(y = stat(y)),
                    
                    required_aes = c("x")
)
```


```{r}
filter_graph_data <- function(clean_data, group_columns, metric_name){
  graph_data <- clean_data %>% 
    {if(!is.null(group_columns)) group_by_at(., .vars=vars(all_of(group_columns))) else .} %>% 
    mutate(group_households = sum(households)) %>% 
    mutate(group_household_weights = households/group_households) %>% 
    arrange(!!sym(metric_name)) %>% 
    mutate(group_percentile = cumsum(households * group_household_weights),
           overall_percentile = cumsum(households)/sum(households)
           ) %>% 
    ungroup()
  return(graph_data)
}

graph_data <- filter_graph_data(clean_data, group_columns, metric_name)
```


```{r}
grouped_weighted_metrics <- function(graph_data, 
                                         group_columns, 
                                         metric_name, 
                                         metric_cutoff_level, 
                                         upper_quantile_view, 
                                         lower_quantile_view){
  # grouped_weighted_medians <- graph_data %>% 
  #    group_by_at(.vars=vars(all_of(group_columns))) %>% 
  #    summarise(metric_median = if( sum(!is.na(households))<3 ){NA} else { weighted.median(!!sym(metric_name), households, na.rm=TRUE)})
  
  weighted_metrics <- graph_data %>% 
    {if(!is.null(group_columns)) group_by_at(., .vars=vars(all_of(group_columns))) else .} %>% 
     summarise(household_count = sum(households),
               total_na = sum(is.na(!!sym(metric_name)) * households, na.rm = TRUE), 
               households_below_cutoff = sum((!!sym(metric_name) < metric_cutoff_level) * households, na.rm = TRUE), 
               metric_max = max(!!sym(metric_name), na.rm = TRUE),
               metric_min = min(!!sym(metric_name), na.rm = TRUE),
               metric_mean = weighted.mean(x=!!sym(metric_name), w=households, na.rm = TRUE),  
               metric_median = if( sum(!is.na(households))<3 ){NA} else { weighted.quantile(x=!!sym(metric_name), w=households, probs=c(.5), na.rm=TRUE)},
               metric_upper = if( sum(!is.na(households))<3 ){NA} else { weighted.quantile(x=!!sym(metric_name), w=households, 
                                                                                            probs=c(upper_quantile_view), na.rm=TRUE)},
               metric_lower = if( sum(!is.na(households))<3 ){NA} else { weighted.quantile(x=!!sym(metric_name), w=households, 
                                                                                            probs=c(lower_quantile_view), na.rm=TRUE)}) %>% 
    mutate(households_pct = household_count/sum(household_count),
           pct_in_group_below_cutoff = households_below_cutoff/household_count,
           pct_total_below_cutoff = households_below_cutoff/sum(households_below_cutoff))
  
  # overall_weighted_metrics <- graph_data %>% ungroup() %>%
  #   summarise(metric_median = if( sum(!is.na(households))<3 ){NA} else { weighted.quantile(x=!!sym(metric_name), w=households, probs=c(.5), na.rm=TRUE)},
  #             metric_upper = if( sum(!is.na(households))<3 ){NA} else { weighted.quantile(x=!!sym(metric_name), w=households, 
  #                                                                                           probs=c(upper_quantile_view), na.rm=TRUE)},
  #             metric_lower = if( sum(!is.na(households))<3 ){NA} else { weighted.quantile(x=!!sym(metric_name), w=households, 
  #                                                                                           probs=c(lower_quantile_view), na.rm=TRUE)}
  #   )
  
#  all_groups <- as.data.frame(matrix(rep("All",length(group_columns)),nrow=1))
#  names(all_groups) <- group_columns
#  overall_weighted_median <- as_tibble(cbind(all_groups, overall_weighted_median))
  
#  weighted_quantiles <- bind_rows(grouped_weighted_medians,overall_weighted_median) %>% ungroup() %>% 
#    mutate_at(.vars=vars(all_of(group_columns)), .funs=as.factor)
  
#  return(weighted_quantiles)
  return(weighted_metrics)
}

calculate_weighted_metrics <- function(graph_data, 
                                         group_columns, 
                                         metric_name, 
                                         metric_cutoff_level, 
                                         upper_quantile_view, 
                                         lower_quantile_view){
  
  weighted_metrics <- grouped_weighted_metrics(graph_data, 
                                         group_columns=NULL, 
                                         metric_name, 
                                         metric_cutoff_level, 
                                         upper_quantile_view, 
                                         lower_quantile_view)
  
  if(!is.null(group_columns)){
    all_groups <- as.data.frame(matrix(rep("All",length(group_columns)),nrow=1))
    names(all_groups) <- group_columns
    weighted_metrics <- as_tibble(cbind(all_groups, weighted_metrics))
    
    grouped_weighted_metrics <- grouped_weighted_metrics(graph_data, 
                                         group_columns=group_columns, 
                                         metric_name, 
                                         metric_cutoff_level, 
                                         upper_quantile_view, 
                                         lower_quantile_view)
    
    weighted_metrics <- bind_rows(grouped_weighted_metrics,weighted_metrics) %>% ungroup() %>% 
      mutate_at(.vars=vars(all_of(group_columns)), .funs=as.factor)
    
    
  }else{
    weighted_metrics <- data.frame(group=as.factor(rep("All", nrow(weighted_metrics))), weighted_metrics)
  }
  
  return(weighted_metrics)
                                         }

# weighted_metrics <- calculate_weighted_metrics(graph_data, 
#                                          group_columns, 
#                                          metric_name, 
#                                          metric_cutoff_level, 
#                                          upper_quantile_view, 
#                                          lower_quantile_view)
# weighted_metrics
```


```{r}
density_chart <- function(graph_data, 
                          metric_name, 
                          metric_label,
                          group_columns, 
                          metric_cutoff_level, 
                          metric_cutoff_label, 
                          chart_title, 
                          chart_subtitle){
  
  if(!is.null(group_columns)){
    pal_n <- length(levels(interaction(graph_data[,group_columns])))
  } else {
    pal_n <- 1
  }
  
  movie <- "Darjeeling1" #"GrandBudapest1"
  #pal <- wes_palette(name=movie, n=pal_n, type="continuous")
  pal <- sample(x=wes_palette(name=movie, n=pal_n, type="continuous"), 
                size = pal_n, 
                replace = FALSE)
  
  weighted_metrics <- calculate_weighted_metrics(graph_data, 
                                         group_columns, 
                                         metric_name, 
                                         metric_cutoff_level, 
                                         upper_quantile_view, 
                                         lower_quantile_view)
  
  
  chart <- graph_data %>%
    # ggplot(aes(x=!!sym(metric_name), 
    #            weight=group_household_weights,
    #            color=interaction(!!!sym(group_columns)),
    #            fill=interaction(!!!sym(group_columns))
    # )) + 
    ggplot(aes_string(x=metric_name, 
               weight="group_household_weights",
               color=if(is.null(group_columns)){group_columns}else{
                 #interaction(!!!sym(group_columns))
                 paste0("interaction(", paste0(group_columns, collapse =  ", "), ")")
                 },
               fill=if(is.null(group_columns)){group_columns}else{
                 #interaction(!!!sym(group_columns))
                 paste0("interaction(", paste0(group_columns, collapse =  ", "), ")")
                 },
               linetype=if(is.null(group_columns)){group_columns}else{
                 #interaction(!!!sym(group_columns))
                 paste0("interaction(", paste0(group_columns, collapse =  ", "), ")")
                 }
    )) + 
    stat_ewcdf(geom='line',  alpha=1) + 
    stat_ewcdf(aes(ymin=..y.., ymax=1), geom='ribbon', alpha=.1) + 
    theme_minimal() + 
    scale_color_manual(values=pal) + 
    scale_fill_manual(values=pal) + 
    scale_x_continuous(labels = scales::dollar_format(accuracy=1),
                       breaks=seq(from=0,to=100,by=10), 
                       minor_breaks=seq(from=0,to=100,by=5),
                       name=metric_label) + 
    scale_y_continuous(labels = scales::label_percent(accuracy = 1), 
                       breaks=seq(from=0,to=1,by=.1), 
                       minor_breaks=seq(from=0,to=1,by=.05),
                       name="Proportion of Households") + 
    theme(legend.justification = c(1, 1), 
          legend.position = c(0.25, 1), 
          legend.title=element_blank(),
          panel.background = element_rect(fill="#f1f1f1"),
          panel.grid.major = element_line(color="#DCDCDC"),
          panel.grid.minor = element_line(color="#DCDCDC"),
          axis.line = element_line(color = "black",
                                   size = 0.5, 
                                   linetype = "solid"),
          axis.text.x=element_text(angle=45, 
                                   hjust=1,
                                   vjust=NULL,
                                   margin=margin(t = 5, 
                                                 r = 0, 
                                                 b = 0, 
                                                 l = 0, 
                                                 unit = "pt")),
          axis.text.y=element_text(angle=10, 
                                   hjust=1,
                                   vjust=0.5,
                                   margin=margin(t = 0, 
                                                 r = 5, 
                                                 b = 0, 
                                                 l = 0, 
                                                 unit = "pt")),
          axis.ticks=element_line(color = "black"),
          axis.ticks.length = unit(-0.1, "cm")) + 
    # geom_segment(y = 0,
    #              x = as.numeric(weighted_medians[weighted_medians$group=="All",c("median_eroi")]),
    #              yend = 0.5,
    #              xend = as.numeric(weighted_medians[weighted_medians$group=="All",c("median_eroi")]),
    #              color="gray25",
    #              linetype="dotted",
    #              size=0.25,
    #              alpha=0.5) +
    # geom_segment(y = 0.5,
    #              x = as.numeric(weighted_medians[weighted_medians$group=="All",c("median_eroi")]),
    #              yend = 0.5,
    #              xend = 0,
    #              color="gray25",
    #              linetype="dotted",
    #              size=0.25,
    #              alpha=0.5) +
    geom_vline(xintercept = metric_cutoff_level,
               linetype="dotted",
               color = "red",
               size=1.0,
               alpha=0.75) +
    annotate("text",
             y = 0,
             x = metric_cutoff_level,
             angle = 0,
             color="red",
             label = metric_cutoff_label,
             vjust = -0.5,
             hjust = 0.0,
             parse = FALSE,
             alpha=0.75) +
    # annotate("text", 
    #          y = 0, 
    #          x = max(weighted_medians$median_eroi), 
    #          angle = 0, 
    #          color="gray25", 
    #          label = "Median", 
    #          vjust = -0.25, 
    #          hjust = -0.1, 
    #          parse = FALSE, 
    #          alpha=0.75) + 
    labs(
      title=chart_title,
      subtitle=chart_subtitle,
      caption=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("By ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          }) + 
    coord_flip(xlim=c(0,100),
               ylim=c(0,1),
               expand=FALSE)
  
  return(chart)
}
```


```{r}
make_all_charts <- function(clean_data,
                            group_columns,
                            metric_name,
                            metric_label, 
                            metric_cutoff_level,
                            metric_cutoff_label,
                            upper_quantile_view,
                            lower_quantile_view,
                            chart_title,
                            chart_subtitle){
  graph_data <- filter_graph_data(clean_data, group_columns, metric_name)
  
  weighted_metrics <- calculate_weighted_metrics(graph_data, 
                                                 group_columns, 
                                                 metric_name, 
                                                 metric_cutoff_level, 
                                                 upper_quantile_view, 
                                                 lower_quantile_view)
  
  density_chart <- density_chart(graph_data, 
                                 metric_name, 
                                 metric_label, 
                                 group_columns, 
                                 metric_cutoff_level, 
                                 metric_cutoff_label, 
                                 chart_title, 
                                 chart_subtitle)
  
  return(density_chart)
}
```



We are primarily interested in how households of different characteristics are distributed according to their Net Energy Ratio, which represents how many net dollars are earned by a household for every dollar they spend on energy. In this context we will refer to NER as the "Net Energy Income Ratio", or the "Net Household Income Earned for Every Dollar Spent on Energy."


```{r}
graph_data <- filter_graph_data(clean_data=clean_data, 
                                group_columns=group_columns, 
                                metric_name=metric_name)
density_chart(graph_data=graph_data,
                group_columns=group_columns,
                metric_name=metric_name, 
              metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```


```{r}
group_columns <- c("in_poverty")

make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```


```{r}
group_columns=c("income_bracket")
make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r results='asis'}
cross_tab <- calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F,
                            bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
cross_tab
```

```{r}
group_columns=c("primary_heating_fuel")
make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```

```{r}
clean_data$simplified_primary_heating_fuel <- as.factor(ifelse(clean_data$primary_heating_fuel=="ELECTRICITY","ELECTRICITY", 
                                                               ifelse(clean_data$primary_heating_fuel %in% c("UTILITY GAS","BOTTLED GAS"), "GAS", 
                                                                      "OTHER")))
group_columns=c("simplified_primary_heating_fuel")

make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```
```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```

```{r}
clean_data$solar_primary_heating_fuel <- as.factor(ifelse(clean_data$primary_heating_fuel=="ELECTRICITY","ELECTRICITY", 
                                                               ifelse(clean_data$primary_heating_fuel %in% c("UTILITY GAS","BOTTLED GAS"), "GAS", 
                                                                      ifelse(clean_data$primary_heating_fuel=="SOLAR","SOLAR",
                                                                      "OTHER"))))
group_columns=c("solar_primary_heating_fuel")

make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```

```{r}
group_columns=c("solar_primary_heating_fuel", "in_poverty")

make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```


```{r}
group_columns=c("occupancy_type")
make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```
```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```

```{r}
group_columns=c("number_of_units")
make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```

```{r}
group_columns=c("year_constructed")
make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```

```{r}
group_columns=c("state_code_usps")
make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```

```{r eval=FALSE}
group_columns=c("occupancy_type")
make_all_charts(clean_data=clean_data,
                group_columns=group_columns,
                metric_name=metric_name, 
                metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                upper_quantile_view=upper_quantile_view,
                lower_quantile_view=lower_quantile_view,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)
```

```{r eval=FALSE, results='asis'}
calculate_weighted_metrics(graph_data=filter_graph_data(clean_data, group_columns, metric_name), 
                                               group_columns, 
                                               metric_name, 
                                               metric_cutoff_level, 
                                               upper_quantile_view, 
                                               lower_quantile_view) %>% t() %>% 
  kable() %>% kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T, border_right = T) %>%
  footnote(general=if(is.null(group_columns)){
        group_columns
        } else {
          paste0("Grouped By: ",paste(group_columns,
                             sep="_",
                             collapse="+"))
          })
```

```{r eval=FALSE}
cliff_chart <- function(){}

ggplot(data=clean_data, aes(x=eroi, y=dear, fill=income_bracket, color=income_bracket)) + 
  xlim(c(0,100)) + 
  ylim(c(0,1)) + 
  geom_point(aes(size=households), alpha=0.1, position="jitter") + 
  geom_hline(yintercept = .9) + 
  geom_vline(xintercept = 10) + 
  geom_smooth(method="lm",
              se=FALSE,
              fullrange=TRUE)
```





```{r eval=FALSE}
scatter_chart <- function(){}

ggplot(data=clean_data, aes(x=annual_income, y=mean_energy_cost, color=income_bracket)) + 
  xlim(c(0,200000)) +
  geom_point(aes(size=households), alpha=0.1) + 
  geom_smooth(method=lm,
                se=FALSE,
                fullrange=TRUE)
```
