---
title: 'PLAN 672: Final Project Report'
author: "Eric Scheier"
date: "`r format(Sys.time(), '%Y-%B-%d')`"
output:
  tufte::tufte_handout: default
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
is_final=FALSE
is_preview=TRUE
is_draft=TRUE
set.seed(123)

knitr::opts_chunk$set(comment='##', 
                      collapse=ifelse(is_preview,TRUE,!is_draft),
                      echo=ifelse(is_preview,FALSE,is_draft),
                      eval=TRUE,
                      warning=ifelse(is_preview,FALSE,is_draft),
                      error=ifelse(is_preview,FALSE,is_draft),
                      results=ifelse(is_final,'hide',ifelse(is_preview,'hide','asis')),
                      fig.keep='all',
                      message=ifelse(is_preview,FALSE,is_draft),
                      include=ifelse(is_preview,TRUE,is_draft),
                      tidy=TRUE,
                      cache=FALSE,
                      fig.margin=FALSE,
                      fig.fullwidth = FALSE
                      )
```

```{r include=FALSE}
library(tidyverse)
library(knitr)
kable(knitr::opts_chunk$get() %>% enframe())
```


```{r}
library(sf)
library(MASS)
library(scales)
library(caret)
library(tidyverse)
library(spatstat)
library(GGally)
library(ggpubr)
library(patchwork)
library(mgcv)
```

```{r}
data <- read_csv(file.path(getwd(),"clean_lead.csv"),
                          col_types=cols(
                            geo_id = readr::col_character(),
                            state_id = readr::col_character(),
                            county_id = readr::col_character(),
                            tract_id = readr::col_character(),
                            puma10_code = readr::col_character(),
                            fmr_code = readr::col_character(),
                            occupancy_type = readr::col_factor(),
                            income_bracket = readr::col_factor(),
                            primary_heating_fuel = readr::col_factor(),
                            number_of_units = readr::col_factor(),
                            year_constructed = readr::col_factor(),
                            households = col_double(),
                            annual_income = col_double(),
                            electricity_spend = col_double(),
                            gas_spend = col_double(),
                            other_spend = col_double(),
                            acs_responses = col_double(),
                            min_age = col_double(),
                            min_units = col_double(),
                            detached = readr::col_factor(NULL),
                            mean_energy_cost = col_double(),
                            mean_energy_burden = col_double(),
                            mean_eroi = col_double(),
                            energy_poverty = col_logical(),
                            pct_non_male = col_double(), 
                            median_age = col_double(),
                            pct_no_high_school = col_double(),
                            pct_non_white = col_double(),
                            median_income = col_double(),
                            acs_total_households = col_double(),
                            utility_id = readr::col_factor(levels = NULL, ordered = FALSE, include_na = TRUE),
                            percent_tract_served = readr::col_double(),
                            tract_households_served = readr::col_double(),
                            utility_name = readr::col_character(),
                            ownership_type = readr::col_factor(NULL),
                            balancing_authority = readr::col_factor(NULL),
                            retail_revenue_thousands = readr::col_double(),
                            retail_sales_MWh = readr::col_double(),
                            retail_customers = readr::col_double(),
                            implied_price_kWh = readr::col_double(),
                            revenue_per_customer = readr::col_double(),
                            implied_electricity_use = readr::col_double()
                            ))
clean_data <- drop_na(data)
```


# Introduction

Energy burden is a statistic defined as the proportion of household income spent on energy bills.  The procurement of electricity, natural gas, fuel oil, wood, coal, solar, and other fuel sources included in this metric are considered essential to the health and safety of prosperous societies. The lower the energy burden, the more discretionary income is available to the household for other necessary goods and services, for savings and investments that contribute to economic growth and community well-being. Inversely, high energy burdens constrict households’ participation in society, and are a contributing factor in poverty broadly.  While per-unit energy costs in the United States are lower than many other countries on average, many Americans still struggle to afford their energy bills.  The universal and fundamental nature of this metric makes it one of public policy interest.  The Federal Low Income Home Energy Assistance Program (LIHEAP) and the Weatherization Assistance Program (WAP) seek to address aspects of energy poverty through bill payment assistance and energy efficiency measures. 

As part of these programs, energy burden has been assessed comprehensively by census tract, county, state, and nationally.  Research based on this data and related methods have found that gender, age, housing age, tenure type, energy inefficiency, education, employment, geography, socioeconomic status, and race/ethnicity are associated with high energy burdens.  However, energy burden has not been examined comprehensively at the scope of the energy provider service territory.  

The energy service territory is more relevant than these other cadastral scopes because the service provider and its regulatory oversight are the authorities for determining the costs of energy for each of its customers: in vertically integrated energy markets the monopoly utility is the only option available to all consumers, and in organized energy markets the utility is designated as the provider of last resort for those unable or unwilling to participate in competitive procurement of energy. In some markets, specialized rates or programs are available for Low and Moderate Income consumers (LMI), who may have higher energy burdens.  In fact, 51% of all funding directed to address high energy burdens in the U.S. is from utility ratepayer funded bill and energy efficiency assistance^[“Lifting the High Energy Burden in America's Largest Cities: How Energy Efficiency Can Improve Low Income and Underserved Communities,” Energy Efficiency for All. [Online]. Available: https://www.energyefficiencyforall.org/resources/lifting-the-high-energy-burden-in-americas-largest-cities-how-energy/. [Accessed: 27-Apr-2020].]. This relevance especially holds true for electricity because it is a commodity delivered via a stationary grid system. Even in organized markets where energy supply is competitive, local utilities retain ownership of the transmission and distribution of electricity to end consumers. Furthermore, the lack of storage infrastructure on the grid and behind each meter means that households are beholden to electricity providers for the time of use.  It stands to reason that energy costs, and therefore energy burdens, might be related to the electric service territory in which a household is located.

Furthermore, the statistic of energy burden is fundamentally problematic because it is a nonlinear transformation in which the denominator (income) may be close to zero. This causes households which have normal energy costs and very low incomes to have high energy burdens. However, the inverse of energy burden is "energy return on investment" (EROI) which is common in the study of energy economics. EROI represents the energy returned from a process such as oil drilling or solar electricity generation. EROI is a unitless factor derived from the numerator and denominator dictated in monetary (e.g. $ returned per $ invested) or energy units (e.g. kWh returned per kWh invested): return / investment. In the context of this study, EROI can represent the economic return that a household receives for every dollar they invest in powering the home (income / energy expenditures). Since the income is in the numerator and energy expenditures are in the denominator of the EROI equation, it avoids the drastically high outliers present in energy burden and is therefore a more suitable metric for assessing the same concept.

This project will estimate the dynamics of household energy return on investment across each electric service territory in North Carolina in order to understand whether household energy returns on investment vary by the type of retail electricity service providers. 

# Data

This analysis relies on two primary sources of data: Low-Income Energy Affordability Data ([LEAD](https://openei.org/doe-opendata/dataset/celica-data)) furnished by the U.S.  Department of Energy and the Electric Retail Service Territories ([ERST](https://hifld-geoplatform.opendata.arcgis.com/datasets/c4fd0b01c2544a2f83440dab292f0980_0)) provided by the U.S. Department of Homeland Security.  It will also rely on two datasets from which the LEAD dataset was formed: Energy Information Agency Form 861 ([EIA](https://www.census.gov/data/developers/data-sets/acs-5year.2016.html)) and the 5-Year American Community Survey ([ACS](https://www.eia.gov/electricity/data/eia861/)). All datasets are openly available and were collected in or for 2016.

+ *Low-Income Energy Affordability Data (LEAD)*
++ Source: Department of Energy
++ Provides: Energy Cost by Housing Type
++ Form: Spreadsheet by Census Tract

+ *Electric Retail Service Territories (ERST)*
++ Source: Department of Homeland Security
++ Provides: Extent of Electricity Grids
++ Form: Shapefiles by Utility

+ *Energy Information Agency Form 861 (EIA)*
++ Source: Energy Information Agency
++ Provides: Revenues, Sales, and Customers
++ Form: Spreadsheet by Utility

+ *5-Year American Community Survey (ACS)*
++ Source: Census Bureau
++ Provides: Demographic Data
++ Form: Shapefiles by Census Block/Tract

As part of the calibration process in the creation of the LEAD dataset, utility service territories were mapped to census tracts using a simple intersection.  The average of energy prices was used for tracts with more than one utility present.  The consumption and expenditure estimates for each tract were rescaled and weighted based on these assumptions. This process means that the consumption and cost assumptions along census tracts shared by multiple utilities (i.e. along territory borders) may show a mean-reversion bias.  These assumptions could be relaxed by recreating the LEAD dataset from scratch from the underlying data and methods while incorporating the spatial distribution of housing units in each tract by block group, but this is outside the scope of this project. Instead, a future analysis will assess the systemic bias of the dataset along service territory borders to determine whether this is an issue. 
Lower income groups may have variable or seasonal income that might not show up in the annualized census data.  These monthly peaks are more important to households when energy bills are higher (e.g. summer air conditioning or winter heating). Furthermore, the impacts of rate structures are not accounted for in the consumption estimates.  This dataset does not control for weather, so the relative efficiency of households cannot be inferred. 


# Methods

I cleaned the [LEAD]() and [ERST]() data according to the process outlined in the linked data munging scripts. In summary, this process involved:

+ *Trivial Steps*
++ Join LEAD ↔ ACS by GEOID
++ Join ERST ↔ EIA by Utility ID
++ Compute row-wise statistics
+ *Non-Trivial Steps*
++ Remove Utility overlap
+++ Filter out incorrect extents
+++ Split overlaps “in half”
+++ Prioritize merge by type
++++ Muni > Coop > IOU
++ Assign Tracts to Utilities
+++ Assign Block Groups by Most Area covered by Utility
+++ Assign Tract by Most Homes Served by Utility

Then, I joined the census tract demographic data, including the LEAD and ACS data, to the information about each electric retail service provider, including the EIA data. This resulted in the following map of census tracts colored by type of service provider serving them:

```{r}
tract_utilites <- st_read("tract_utilities.geojson")

tract_utilites %>% 
  ggplot() +
  geom_sf(aes(fill=ownership_type),size=.1) + 
  coord_sf(crs = 4326) + 
  theme(legend.position = "top")
```


# Results

This process resulted in a dataset which outlines the energy expenditures, incomes, household characteristics, and demographic characteristics for each census tract in North Carolina.

```{r}
str(clean_data) #print(kable(var_desc) %>% kable_styling(c("striped", "bordered")))
summary(clean_data)
```

```{r}
weighted_medians <- clean_data %>%
   group_by(ownership_type) %>% 
   summarise(median_electricity_spend = if( sum(!is.na(households))<3 ){NA} else { weighted.median(electricity_spend, households, na.rm=TRUE)},
             median_eroi = if( sum(!is.na(households))<3 ){NA} else { weighted.median(mean_eroi, households, na.rm=TRUE)},
            median_income = if( sum(!is.na(households))<3 ){NA} else { weighted.median(annual_income, households, na.rm=TRUE)},
            median_electricity_price = if( sum(!is.na(households))<3 ){NA} else { weighted.median(implied_price_kWh, households, na.rm=TRUE)},
            median_electricity_use = if( sum(!is.na(households))<3 ){NA} else { weighted.median(implied_electricity_use, households, na.rm=TRUE)},
            median_energy_burden = if( sum(!is.na(households))<3 ){NA} else { weighted.median(mean_energy_burden, households, na.rm=TRUE)},
            median_energy_cost = if( sum(!is.na(households))<3 ){NA} else { weighted.median(mean_energy_cost, households, na.rm=TRUE)}
            )

clean_data <- clean_data %>% 
  group_by(ownership_type) %>% 
  mutate(ownership_households = sum(households)) %>% 
  mutate(ownership_type_household_weights = households/ownership_households)

weighted_medians
```

## Energy Return on Investment (EROI)

The household energy return on investment is calculated from this dataset by dividing annual household income by annual household energy expenditures. Energy poverty is commonly defined as an expenditure of greater than 10% of household income on energy^[D. J. Bednar and T. G. Reames, “Recognition of and response to energy poverty in the United States,” Nature Energy, 2020.], which translates to an EROI of less than 10x.

```{r}
# EROI
mean_eroi <- clean_data %>%
  ggplot(aes(x=mean_eroi, weight=ownership_type_household_weights/3, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Household Energy Return on Investment By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::unit_format(unit = "x", scale = 1, accuracy = 1),
                     breaks=seq(from=0,to=100,by=10), 
                     minor_breaks=seq(from=0,to=100,by=1),
                     limits=c(0,100), name="Household Energy Return on Investment\n(Income Earned for each Dollar Spent on Energy)") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.05), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_eroi,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75) + 
  geom_vline(xintercept = 10, linetype="dotted", 
                color = "red", size=1.0, alpha=0.75)

mean_eroi + annotate("text", x = 10, y = 0.0075/3, angle = 90, color="red", label = "Energy Poverty Line", 
    vjust = -0.5, parse = FALSE, alpha=0.75) + 
  annotate("text", x = min(weighted_medians$median_eroi), y = 0.005/3, angle = 90, color="gray25", label = "Median", 
    vjust = -0.5, parse = FALSE, alpha=0.75)
```

It seems that energy return on investment assumes a bimodal distribution, and that households served by municipal electric utilities have a significantly lower EROI than those served by cooperatives and investor owned utilities. Municipal electricity providers have more of their households in the 5x-20x range, and more in the 40x-60x range, while cooperatives and investor-owned utilities have a fatter right tail of high-EROI customers (60x+).

What explains these distributions? Examining the component parts of EROI reveals the source of the bimodal distribution and fat tails.

## Annual Income

The annual incomes of households in North Carolina appears to be bimodal.

```{r}
# Annual Income
federal_poverty_line <- 16240
annual_income <- clean_data %>% #drop_na() %>%
  ggplot(aes(x=annual_income, weight=ownership_type_household_weights/3, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Household Annual Income By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::dollar_format(), limits=c(0,150000), name="Annual Income") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.0005), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_income,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75) +
  geom_vline(xintercept = federal_poverty_line, linetype="dotted",
                color = "red", size=1.0, alpha=0.75)

annual_income +
  annotate("text", x = federal_poverty_line, y = 0.0000035/3, angle = 90, color="red", label = "Federal Poverty Line",
    vjust = -0.5, parse = FALSE, alpha=0.75) +
  annotate("text", x = min(weighted_medians$median_income), y = 0.0000025/3, angle = 90, color="gray25", label = "Median",
    vjust = -0.5, parse = FALSE, alpha=0.75)
```


## Monthly Energy Expenditures

The monthly expenditures of households in North Carolina seem to complete the picture seen in the EROI chart. Investor owned utility customers spend less on energy, and cooperative customers also spend slightly less. Municipal customers spend the most, including a significantly larger portion of the customer base spending more than $180/month on energy. 

```{r}
# Monthly Energy Expenditures
mean_energy_cost <- clean_data %>%
  ggplot(aes(x=mean_energy_cost, weight=ownership_type_household_weights/3, 
             fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Household Monthly Energy Spending By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::dollar_format(), limits=c(0,300), name="Energy Spend",
                     breaks=seq(from=0,to=300,by=50), 
                     minor_breaks=seq(from=0,to=300,by=10)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.0005), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_energy_cost,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75)

mean_energy_cost + annotate("text", x = min(weighted_medians$median_energy_cost), 
                            y = 0.0025/3, angle = 90, color="gray25", 
                            label = "Median",vjust = -0.5, parse = FALSE, alpha=0.75)
```

Together, these factors show that municipal electricity customers earn less per household for every dollar invested in powering the household than their counterparts in cooperative and investor owned service territories.

```{r}
# Annual Income vs Monthly Energy Expenditures
# https://stackoverflow.com/a/59581217/2668097

energy_v_income_contour_plot <- ggplot(clean_data, aes(x=mean_energy_cost,
                                       y=annual_income,
                                       weight=ownership_type_household_weights/3,
                                       color=ownership_type,
                                       fill=ownership_type)) +
  stat_density_2d(aes(alpha = ..piece..), geom="polygon") +
  guides(alpha = FALSE) +
  stat_smooth(method = "lm", fullrange = TRUE) +
  geom_rug() + 
  scale_x_continuous(name = "Monthly Energy Cost", 
                     labels = scales::dollar_format(),
                     limits = c(0, 300), expand = c(0, 0)) + 
  scale_y_continuous(name = "Annual Income", 
                     labels = scales::dollar_format(),
                     limits = c(0, 150000), expand = c(0, 0)) + 
  theme_pubr() + #   theme(plot.margin = margin()) + 
  theme(legend.position = "bottom")

mean_energy_cost + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  ggtitle("Monthly Energy Cost vs. Annual Income") + 
  plot_spacer() + 
  energy_v_income_contour_plot + theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  annual_income + coord_flip() + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
```

# Electricity

However, this analysis focuses on electricity, which composes only a portion of total energy expenditures. How does electricity consumption specifically contribute to the difference among EROI by service provider type?

## Electricity Expenditures

Apparently, customers of investor owned electric utilities spend approximately $10 less per month on electricity than their counterparts.

```{r}
# Monthly Electricity Expenditures
electricity_spend <- clean_data %>%
  ggplot(aes(x=electricity_spend, weight=ownership_type_household_weights/3, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Household Monthly Electricity Spending By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::dollar_format(), limits=c(0,300), name="Energy Spend",
                     breaks=seq(from=0,to=300,by=50), 
                     minor_breaks=seq(from=0,to=300,by=10)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.0005), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_electricity_spend,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75)

electricity_spend + annotate("text", x = min(weighted_medians$median_electricity_spend), y = 0.0025/3, angle = 90, color="gray25", label = "Median",vjust = -0.5, parse = FALSE, alpha=0.75)
```

However, cooperative customers are the only ones for whom the relationship between EROI and electricity expenditures is positive. This implies that cooperative customers are using electricity to improve their welfare more than their counterparts.

```{r}
# EROI vs Monthly Electricity Expenditures
electricity_v_eroi_contour_plot <- ggplot(clean_data, aes(x=electricity_spend,
                                       y=mean_eroi,
                                       weight=ownership_type_household_weights/3,
                                       color=ownership_type,
                                       fill=ownership_type)) +
  stat_density_2d(aes(alpha = ..piece..), geom="polygon") +
  guides(alpha = FALSE) +
  stat_smooth(method = "lm", fullrange = TRUE) +
  geom_rug() + 
  scale_x_continuous(name = "Monthly Electricity Cost", 
                     labels = scales::dollar_format(),
                     limits = c(0, 300), expand = c(0, 0)) + 
  scale_y_continuous(name = "EROI", 
                     labels = scales::unit_format(unit = "x", scale = 1, accuracy = 1),
                     limits = c(0, 100), expand = c(0, 0)) + 
  theme_pubr() + #   theme(plot.margin = margin()) + 
  theme(legend.position = "bottom")

electricity_spend + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  ggtitle("Monthly Electricity Cost vs. EROI") + 
  plot_spacer() + 
  electricity_v_eroi_contour_plot + theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  mean_eroi + coord_flip() + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
```


## Electricity Rates

Is this behavior explained by rates? Partially: investor owned utilities offer significantly lower rates than cooperatives, which in turn offer double the discount compared to municipal service providers. This helps explain why investor owned utility customers spend less monthly on electricity, and implies that cooperative customers use more electricity than municipal customers (for them to be spending the same monthly amount). This makes sense, as cooperative customers would be rational to spend more on electricity if it is providing them with a positive marginal EROI.


```{r}
# Electricity Rates
implied_price_kWh <- clean_data %>%
  ggplot(aes(x=implied_price_kWh, weight=ownership_type_household_weights/3000, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Electricity Price By Utility Ownership Type") + 
  scale_x_continuous(labels = scales::dollar_format(), limits=c(0.1,0.15), name="Electricity Unit Cost ($/kWh)",
                     breaks=seq(from=0.1,to=0.15,by=0.01), 
                     minor_breaks=seq(from=0.1,to=0.15,by=0.0025)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_electricity_price,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75)

implied_price_kWh +
  annotate("text", x = min(weighted_medians$median_electricity_price), y = 0.1, angle = 90, color="gray25", label = "Median",vjust = -0.5, parse = FALSE, alpha=0.75)
```

Household EROI seems to diminish quickly with increases in electricity rates among investor owned utilities, whereas the elasticity of EROI with respect to price is low cooperative members and lower for municipal customers.

```{r}
# EROI vs Electricity Rates
rates_v_eroi_contour_plot <- ggplot(clean_data, aes(x=implied_price_kWh,
                                       y=mean_eroi,
                                       weight=ownership_type_household_weights/3,
                                       color=ownership_type,
                                       fill=ownership_type)) +
  stat_density_2d(aes(alpha = ..piece..), geom="polygon") +
  guides(alpha = FALSE) +
  stat_smooth(method = "lm", fullrange = TRUE) +
  geom_rug() + 
  scale_x_continuous(name = "Electricity Unit Cost ($/kWh)", 
                     labels = scales::dollar_format(),
                     limits = c(0.1, 0.15), expand = c(0, 0)) + 
  scale_y_continuous(name = "EROI", 
                     labels = scales::unit_format(unit = "x", scale = 1, accuracy = 1),
                     limits = c(0, 100), expand = c(0, 0)) + 
  theme_pubr() + #   theme(plot.margin = margin()) + 
  theme(legend.position = "bottom")

implied_price_kWh + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  ggtitle("Electricity Rates vs. EROI") + 
  plot_spacer() + 
  rates_v_eroi_contour_plot + theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  mean_eroi + coord_flip() + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
```

## Electricity Consumption

The implied electricity consumption for each cohort can be estimated based on the average rate for the utility serving that cohort and esitmated electricity expenditures. This estimate should be compared to a more direct survey of energy consumption and the results of other estimation procedures, but can serve as a useful metric in the meantime.

While cooperative and municipal customer usage appears to be normally distributed, investor owned utility usage appears diminished for households consuming 1-1.5 MWh/month. This could be explained by energy efficiency programs, or consumers above certain levels of consumption substituting away from electricity.

This estimate shows that cooperative customers do use more electricity than municipal customers, and that investor owned utility customer in turn use more than both of their counterparts. This makes sense given the relative pricing and EROI tradeoffs of these respective options shown previously.

```{r}
# Implied Electricity Consumption
implied_electricity_use <- clean_data %>%
  ggplot(aes(x=implied_electricity_use, weight=ownership_type_household_weights/3, fill=ownership_type, color=ownership_type)) + 
  geom_density(alpha=0.1) + 
  ggtitle("Implied Monthly Household Electricity Use By Utility Ownership Type") + 
  scale_x_continuous(labels = comma, limits=c(0,2200), 
                     name="Electricity Consumption (kWh)",
                     breaks=seq(from=0,to=2200,by=500), 
                     minor_breaks=seq(from=0,to=2200,by=100)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.0005), 
                     name="Proportion of Households") + 
  theme_minimal() + 
  theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  geom_vline(data=weighted_medians, aes(xintercept=median_electricity_use,  color=ownership_type),
               linetype="solid", size=0.5, alpha=0.75)

implied_electricity_use +
  annotate("text", x = min(weighted_medians$median_electricity_use), y = 0.00015, angle = 90, color="gray25", label = "Median",vjust = -0.5, parse = FALSE, alpha=0.75)
```

The comparison of electricity consumption to EROI does not provide additional information to the same graphic comparison of EROI to electricity expenditures, other than to say that the positive EROI relationship for cooperative customers and slight negative relationship for municipal and investor owned utility customers holds up when adjusted for the relative prices offered by each provider.

```{r}
# EROI vs Implied Electricity Consumption
electricity_use_v_eroi_contour_plot <- ggplot(clean_data, aes(x=implied_electricity_use,
                                       y=mean_eroi,
                                       weight=ownership_type_household_weights/3,
                                       color=ownership_type,
                                       fill=ownership_type)) +
  stat_density_2d(aes(alpha = ..piece..), geom="polygon") +
  guides(alpha = FALSE) +
  stat_smooth(method = "lm", fullrange = TRUE) +
  geom_rug() + 
  scale_x_continuous(name = "Implied Monthly Household Electricity Use (kWh)", 
                     labels = comma,
                     limits = c(0, 2200), expand = c(0, 0)) + 
  scale_y_continuous(name = "EROI", 
                     labels = scales::unit_format(unit = "x", scale = 1, accuracy = 1),
                     limits = c(0, 100), expand = c(0, 0)) + 
  theme_pubr() + #   theme(plot.margin = margin()) + 
  theme(legend.position = "bottom")

implied_electricity_use + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  ggtitle("Electricity Consumption vs. EROI") + 
  plot_spacer() + 
  electricity_use_v_eroi_contour_plot + theme(legend.justification = c(1, 1), legend.position = c(1, 1), legend.title=element_blank()) + 
  mean_eroi + coord_flip() + theme_void() + theme(legend.position = "none") + labs(title = NULL, x=NULL, y=NULL) + 
  plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
```

In conclusion, it seems that EROIs do vary among electric service provider types in North Carolina. This variance is only partially explained by the price of electricity offered by each provider.

# Confounding Variables

While this exploration of electricity's relationship with household EROI is enlightening, it is not comprehensive. In addition to accounting for other energy consumption (e.g. natural gas, solar power) and other types of consumption (e.g. transportation, food, housing, goods, services), there could be further factors that impact the translation of household energy consumption into wealth. Utilizing the full set of characteristics available in the LEAD and ACS datasets might have explanatory power, and some cursory assessments of the potential relationships are presented below.

## Household Characteristics

A set of household characteristics are collected as part of the LEAD dataset assembly. These include:

+ `occupancy_type`: whether the household is rented or owned.
+ `primary_heating_fuel`: the primary source of heat for the household.
+ `min_age`: the minimum age of the dwelling.
+ `min_units`: the minimum number of living units in the building that contains the household.
+ `detached`: whether the household is attached or detached from neighboring units.

```{r fig.margin=FALSE, fig.fullwidth = TRUE, fig.keep='all', fig.cap='Relationships among household characteristics and EROI'}
# Household Characteristics Pair Plot
household_characteristics <- c(
  "occupancy_type",
  "primary_heating_fuel",
  #"number_of_units",
  #"year_constructed",
  "min_age",
  "min_units",
  "detached",
  "ownership_type",
  "mean_eroi"
)

household_pair_plot <- ggpairs(clean_data, aes(color = ownership_type, weight=ownership_type_household_weights/3, alpha = 0.4),
                     columns = household_characteristics,
                     upper = list(
                       continuous = wrap("cor", size = 1.5, alignPercent = 1)),
                     progress=FALSE) + theme_minimal(base_size = 5) + 
  theme(axis.text=element_text(size=3.0))

ggsave("./household_pair_plot.png", plot=household_pair_plot, width=10,height=7.5,units="in")

suppressMessages(print(household_pair_plot))
```


## Demographic Characteristics

A set of demographic characteristics are calculated from the ACS. These include:

+ `pct_non_male`: The proportion of the census tract population that is not male.
+ `median_age`: The median age of the census tract's population.
+ `pct_no_high_school`: The proportion of the census tract population that has received less than a high school education.
+ `pct_non_white`: The proportion of the census tract population that is not white.
+ `median_income`: The median income of the census tract's population.

```{r fig.cap='Relationships among demographic characteristics and EROI'}
# Demographic Characteristics Pair Plot
demographic_characteristics <- c(
  "pct_non_male",
  "median_age",
  "pct_no_high_school",
  "pct_non_white",
  "median_income",
  "ownership_type",
  "mean_eroi"
)

demographic_pair_plot <- ggpairs(clean_data, aes(color = ownership_type, weight=ownership_type_household_weights/3, alpha = 0.4),
                     columns = demographic_characteristics,
                     upper = list(
                       continuous = wrap("cor", size = 1.5, alignPercent = 1)),
                     progress=FALSE) + theme_minimal(base_size = 5) + 
  theme(axis.text=element_text(size=3.0))

ggsave("./demographic_pair_plot.png", plot=demographic_pair_plot, width=10,height=7.5,units="in")

suppressMessages(print(demographic_pair_plot))
```