---
title: 'Net Energy Equity: Findings'
author: "Eric Scheier"
date: "`r format(Sys.time(), '%Y-%B-%d')`"
output:
  html_document: default
  pdf_document: default
  tufte::tufte_handout: default
  word_document: default
---

```{r setup, include=FALSE}
is_final=FALSE
is_preview=TRUE
is_draft=TRUE
set.seed(123)

knitr::opts_chunk$set(comment='##', 
                      collapse=ifelse(is_preview,TRUE,!is_draft),
                      echo=ifelse(is_preview,FALSE,is_draft),
                      eval=TRUE,
                      warning=ifelse(is_preview,FALSE,is_draft),
                      error=ifelse(is_preview,FALSE,is_draft),
                      results=ifelse(is_final,'hide',ifelse(is_preview,'hide','asis')),
                      fig.keep='all',
                      message=ifelse(is_preview,FALSE,is_draft),
                      include=ifelse(is_preview,TRUE,is_draft),
                      tidy=TRUE,
                      cache=FALSE,
                      fig.margin=FALSE,
                      fig.fullwidth = FALSE
                      )
```

```{r include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
kable(knitr::opts_chunk$get() %>% enframe())
```


```{r}
library(tidyverse)
library(sf)
library(MASS)
library(scales)
library(caret)
library(spatstat)
library(GGally)
library(ggpubr)
library(patchwork)
library(mgcv)
library(tidyselect)
library(RColorBrewer)
library(wesanderson)
library(tigris)
library(readxl)
library(tidyverse)
```


```{r}
source("ratios.R")
source("charts.R")
```

# Loading Data

```{r}
# load the tract or state (or zip) geojson from csv to sf as: data
census_tracts <- st_read("census_tracts.geojson")
states <- st_read("states.geojson")
data <- read_csv("very_clean_data.csv")
```

# Calculating Additional Metrics

## Poverty

How many households are below the poverty line in this data?

```{r}
total_households <- sum(data$households)

number_in_poverty <- sum((data$in_poverty==TRUE) * (data$households))

pct_in_poverty <- number_in_poverty / total_households

number_above_poverty <- sum((data$in_poverty==FALSE) * (data$households))

pct_above_poverty <- number_above_poverty / total_households
```

# Adding Net Energy Analysis Metrics

Most/All Net Energy Analysis metrics are composed of the gross amount of resource extracted, and the amount of resource spent in the extraction process. For households extracting income from the economy, these ratios will be composed of:

`g` = gross income
`s` = spending on energy

How are these metrics related? We are specifically curious about the relationship among households considered to be in poverty by the federal government.

```{r}
ggplot(data=data, aes(x=annual_income, y=annual_energy_cost, color=in_poverty)) + 
  geom_point(aes(size=households), alpha=0.05, show.legend = FALSE) + 
  geom_smooth(method=lm,
                se=TRUE,
                fullrange=TRUE)# + 
  #xlim(0,500000)
  
```

From these metrics we can create all of the relevant energy ratios:

## Traditional Energy Burden Indicator

`energy_burden = s/g`


```{r}
energy_burden_poverty_line <- 0.10

eroi_poverty_line <- eroi_func(g=1,
                               s=energy_burden_poverty_line)

average_energy_cost <- weighted.mean(data$annual_energy_cost, 
                                     data$annual_kWh*data$households, 
                                     na.rm = T)/weighted.mean(data$annual_kWh,
                                                              data$households,
                                                              na.rm = T)

median_energy_cost <- weighted.median(data$annual_energy_cost, 
                                     data$annual_kWh*data$households, 
                                     na.rm = T)/weighted.median(data$annual_kWh,
                                                              data$households,
                                                              na.rm = T)
# 12*(data$replica_electricity_spend + 
#       data$lead_gas_spend + 
#       data$lead_other_spend)
# data$total_kWh <- data$gas_kWh + data$electricity_kWh
median_electricity_cost <- weighted.median(data$replica_electricity_spend, 
                                     data$electricity_kWh*data$households, 
                                     na.rm = T)/weighted.median(data$electricity_kWh,
                                                              data$households,
                                                              na.rm = T)

median_gas_cost <- weighted.median(data$lead_gas_spend, 
                                     data$gas_kWh*data$households, 
                                     na.rm = T)/weighted.median(data$gas_kWh,
                                                              data$households,
                                                              na.rm = T)
median_gas_cost_Mcf <- weighted.median(data$lead_gas_spend, 
                                     data$gas_Mcf*data$households, 
                                     na.rm = T)/weighted.median(data$gas_Mcf,
                                                              data$households,
                                                              na.rm = T)


ner_poverty_line_dlrs <- ner_func(g=1,
                                  s=energy_burden_poverty_line)

ner_poverty_line_mean <- ner_func(g=1,
                             s=energy_burden_poverty_line,
                             se=energy_burden_poverty_line/(average_energy_cost))

ner_poverty_line_median <- ner_func(g=1,
                             s=energy_burden_poverty_line,
                             se=median_energy_cost/energy_burden_poverty_line)

ner_poverty_line <- ner_poverty_line_median


dear_poverty_line <- dear_func(g=1,
                               s=energy_burden_poverty_line)

ner_dear_poverty_line <- dear_func(g=1+median_energy_cost*ner_poverty_line_median,
                               s=1)

```



## Test Run

```{r}
clean_data <- data
```



```{r}
metric_name <- "ner"
metric_cutoff_level <- ner_poverty_line
group_variable <- NULL# "GEOID" #"state_abbr" #merge_geo_id" #
group_columns <- c(group_variable) #c("gisjoin") #
graph_data <- filter_graph_data(data, group_columns, metric_name)

top_metrics <- grouped_weighted_metrics(graph_data, 
                         group_columns, 
                         metric_name, 
                         metric_cutoff_level, 
                         upper_quantile_view=0.999, 
                         lower_quantile_view=0.001)
head(top_metrics)
```


```{r}
#data$GEOID <- sub('.', '', data$gisjoin)
group_variable <- "gisjoin"# "GEOID" #"state_abbr" #merge_geo_id" #
group_columns <- c(group_variable) #c("gisjoin") #
graph_data <- filter_graph_data(clean_data, group_columns, metric_name)

gwm <- grouped_weighted_metrics(graph_data, 
                         group_columns, 
                         metric_name, 
                         metric_cutoff_level, 
                         upper_quantile_view=0.75, 
                         lower_quantile_view=0.25)
head(gwm)
```

```{r}
clean_top_metrics <- grouped_weighted_metrics(graph_data, 
                         group_columns=NULL, 
                         metric_name, 
                         metric_cutoff_level, 
                         upper_quantile_view=0.75, 
                         lower_quantile_view=0.25)
head(clean_top_metrics)
```


```{r}
tract_shp <- left_join(census_tracts, gwm, by=c("gisjoin"))

# cloropleth map by census tract
centroids <- st_coordinates(st_centroid(tract_shp$geometry)) %>% data.frame()
centroids <- centroids[is.finite(rowSums(centroids)),]

# center_centroid <- st_centroid()

o <- tract_shp[,c("metric_median", "geometry")] %>% 
  ggplot() +
  geom_sf(aes(fill=metric_median),size=0, alpha=0.9) + 
  coord_sf(crs = 4326, #26945, #
           xlim = c(min(centroids$X), max(centroids$X)), 
           ylim = c(min(centroids$Y), max(centroids$Y)), 
           expand = TRUE)
o
```


```{r}
color_values <- as.numeric(sort(c(clean_top_metrics$metric_min,
                                clean_top_metrics$metric_lower,
                                clean_top_metrics$metric_median,
                                # top_metrics$pct_in_group_below_cutoff,
                                clean_top_metrics$metric_upper, 
                                clean_top_metrics$metric_max), decreasing = F))
color_values <- scales::rescale(x=color_values, to=c(0,1))

p <- o +
  # geom_polygon(data = spdf_fortified, 
  #              aes(fill = nb_equip, x = long, y = lat, group = group) , 
  #              size=0, alpha=0.9) +
  theme_void() +
  # scale_continuous_identity(aes(color=metric_median)) +
  # scale_fill_identity() +
  scale_fill_gradientn(#) + 
  # scale_fill_viridis(trans = "log", 
                     # breaks=c(1,5,10,20,50,100),
                     name="Net Income per Energy Spend ($/kWh)",
                     colors = c("#ff5252","#fff176","#228b22"),
                     na.value = "#616161",
                     values = color_values,
                     guide = guide_colorbar(direction="horizontal",
                                            title.position = 'top'
                                            ))+#keyheight = unit(4, units = "mm"),
                                           #keywidth=unit(4, units = "mm"),
                                           # label.position = "bottom",
                                           # , nrow=1) ) +
  labs(
    title = "Net Energy Return of U.S. Communities",
    subtitle = "Dollars earned per unit of energy consumed.",
    caption = "Data: NREL" # | Creation: Eric Scheier | emergi.eco"
  ) +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),

    plot.title = element_text(size= 22, 
                              hjust=0.01, 
                              color = "#4e4d47", 
                              margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
    plot.subtitle = element_text(size= 17, 
                                 hjust=0.01, 
                                 color = "#4e4d47", 
                                 margin = margin(b = -0.1, t = 0.43, l = 2, unit = "cm")),
    plot.caption = element_text( size=12, 
                                 color = "#4e4d47", 
                                 margin = margin(b = 0.3, r=-99, unit = "cm") ),

    legend.position = c(.7, .8)
  ) #+
  #coord_map()
p
```

```{r}
tract_shp$in_poverty <- tract_shp$metric_median <= ner_poverty_line
q <- tract_shp[,c("in_poverty", "geometry")] %>% 
  ggplot() +
  geom_sf(aes(fill=in_poverty),size=0, alpha=0.9) + 
  coord_sf(crs = 4326, #26945, #
           xlim = c(min(centroids$X), max(centroids$X)), 
           ylim = c(min(centroids$Y), max(centroids$Y)), 
           expand = TRUE)
q
```

# Analysis

Of the above metrics, the Net Energy Ratio (NER) and Discretionary Energy Availability Rate (DEAR) are the most desirable, The DEAR is nicely bound between 0-100%, while the NER is the standard metric of Net Energy Analysis.

EROI is a precursor of NER, while Energy Burden is the metric du jour in the energy insecurity literature.

```{r eval=FALSE}
ggplot(data=clean_data, aes(x=eroi, y=dear, fill=income_bracket, color=income_bracket)) + 
  xlim(c(0,100)) + 
  ylim(c(0,1)) + 
  geom_point(aes(size=households), alpha=0.1, position="jitter") + 
  geom_hline(yintercept = .9) + 
  geom_vline(xintercept = 10) + 
  geom_smooth(method="lm",
              se=FALSE,
              fullrange=TRUE)
```

DEAR and NER are often compared to portray a phenomenon known as the "net energy cliff" Do we see the energy cliff in these data?

```{r}
ggplot(data=clean_data, aes(x=ner, y=dear, fill=in_poverty, color=in_poverty)) + 
  xlim(c(0,10)) + 
  ylim(c(0,1)) + 
  geom_point(aes(size=households), alpha=0.1, position="jitter", show.legend = FALSE) + 
  geom_hline(yintercept = dear_poverty_line) + 
  geom_vline(xintercept = ner_poverty_line_median) + 
  geom_smooth(method="lm",
              se=FALSE,
              fullrange=TRUE)
```


```{r}
#chart_title <- "Household Economic Return on Energy Spending"
chart_title <- "Community Net Energy Return"
chart_subtitle <- "Net Earnings per Unit of Energy Consumed"

group_columns <- NULL#"income_bracket")#in_poverty
                   #"primary_heating_fuel"

metric_name <- "ner" #"energy_burden" #"ner" #"dear" #"eroi"
metric_label <- "$/kWh"
metric_cutoff_level <- ner_poverty_line
metric_cutoff_label <- "Energy Poverty Line"

upper_quantile_view <- 1.0
lower_quantile_view <- 0.0
```


```{r eval=FALSE}
# for (geo_id in graph_data$merge_geo_id){
group_inspect <- function(geo_id){
    print(as.character(geo_id))
    sub_graph_data <- graph_data[graph_data$gisjoin==geo_id,]
    if( sum(!is.na(sub_graph_data$households))<3 || 
        sum
        all(sub_graph_data$households==0) ){
      print(NA)
      return(FALSE)
    } else {
      return(as.numeric(weighted.quantile(x=sub_graph_data$ner, w=sub_graph_data$households, probs=c(.5), na.rm=T)))
    }
}

sapply(unique(graph_data$gisjoin), group_inspect)
```


We are primarily interested in how households of different characteristics are distributed according to their Net Energy Ratio, which represents how many net dollars are earned by a household for every dollar they spend on energy. In this context we will refer to NER as the "Net Energy Income Ratio", or the "Net Household Income Earned for Every Dollar Spent on Energy."

```{r eval=FALSE}
clean_data$simplified_primary_heating_fuel <- as.factor(ifelse(clean_data$primary_heating_fuel=="ELECTRICITY","ELECTRICITY", 
                                                               ifelse(clean_data$primary_heating_fuel %in% c("UTILITY GAS","BOTTLED GAS"), "GAS", 
                                                                      "OTHER")))
group_columns=c("simplified_primary_heating_fuel")

clean_data$solar_primary_heating_fuel <- as.factor(ifelse(clean_data$primary_heating_fuel=="ELECTRICITY","ELECTRICITY", 
                                                               ifelse(clean_data$primary_heating_fuel %in% c("UTILITY GAS","BOTTLED GAS"), "GAS", 
                                                                      ifelse(clean_data$primary_heating_fuel=="SOLAR","SOLAR",
                                                                      "OTHER"))))
group_columns=c("solar_primary_heating_fuel")
```

```{r}
top_line_group <- "in_poverty"

top_line_charts <- make_all_charts(clean_data,
                            group_columns=top_line_group,
                            metric_name,
                            metric_label, 
                            metric_cutoff_level,
                            metric_cutoff_label,
                            #upper_quantile_view,
                            #lower_quantile_view,
                            chart_title=chart_title,
                            chart_subtitle=chart_subtitle)
top_line_charts
```


```{r}
graph_data <- filter_graph_data(clean_data=clean_data, 
                                group_columns=NULL, 
                                metric_name=metric_name)
density_chart(graph_data=graph_data,
                group_columns=NULL,
                metric_name=metric_name, 
              metric_label=metric_label, 
                metric_cutoff_level=metric_cutoff_level,
                metric_cutoff_label=metric_cutoff_label,
                chart_title=chart_title,
                chart_subtitle=chart_subtitle)

```


```{r}
make_violin_plot <- function(graph_data, 
                             group_columns, 
                             metric_name,
                             metric_cutoff_level,
                             upper_quantile_view=0.99,
                             lower_quantile_view=0.0){
  
  gwm <- grouped_weighted_metrics(graph_data=graph_data, 
                           group_columns=group_columns, 
                           metric_name=metric_name, 
                           metric_cutoff_level=metric_cutoff_level, 
                           upper_quantile_view=upper_quantile_view, 
                           lower_quantile_view=lower_quantile_view)
  
  ylims <- c(min(max(0,gwm$metric_lower[is.finite(gwm$metric_lower)])),
             max(gwm$metric_upper[is.finite(gwm$metric_upper)]))
  
  sort_column <- if(is.null(group_columns)){"gisjoin"}else if(length(group_columns)==1){
    group_columns
  }else{paste0(group_columns[0])}
  
  y <- graph_data %>% #[graph_data$state_abbr=="VT",],
    # mutate(!!sym(group_variable) = fct_reorder(!!sym(group_variable), !!sym(metric_name), .fun='median'))) %>% 
    # mutate(state_abbr = fct_reorder(state_abbr, ner, .fun='median')) %>%
    {if(!is.null(group_columns)) group_by_at(., .vars=vars(all_of(group_columns))) else .} %>% 
    ggplot(aes(group=if(is.null(group_columns)){
      group_columns}else{
        paste0("interaction(", paste0(group_columns, collapse =  ", "), ")")},
      x = fct_reorder(!!sym(sort_column),
                      !!sym(metric_name),
                      .fun=median, na.rm=T, .desc=F), #!!sym(group_variable),#
      y= !!sym(metric_name)))+#, 
               #color = !! sym(response_var), 
               #fill= !! sym(response_var))) + 
      geom_boxplot(aes(weight=sqrt(households),
                      varwidth = T),
                   notch=TRUE, width=0.2, fill=NA, color="gray", shape = 18, size = 0.5,
                   outlier.shape = NA) + 
      stat_summary(fun.y=median,fun.args=c(na.rm=TRUE), 
                   geom = "point", shape = 18, size = 1) + 
      # geom_violin(trim=FALSE, alpha=0.3, outlier.shape=NA) + 
      #ggtitle(paste0("Predictor Scaled Distributions")) +
      theme_minimal(base_size = 5) + theme(legend.position="none") + 
      theme(axis.text=element_text(size=3.5)) + 
    coord_cartesian(ylim=as.numeric(ylims)*1.05)
      # coord_flip(ylim = as.numeric(ylims)*1.05)
  # print("graph_2")
   return(y)
}
graph_data <- filter_graph_data(clean_data=clean_data, 
                                group_columns=c("states"), 
                                metric_name=metric_name)

y <- make_violin_plot(graph_data=graph_data,
                      group_columns=group_columns,
                      metric_name=metric_name,
                      metric_cutoff_level=metric_cutoff_level)
print(y)
```
